---
# title: "Two-Stage Data Practice"
# author: "Jackson Hughes"
date: "4/11/2021"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(skimr)
library(moderndive)
```

### Set working directories

```{r}
proj_dir = here()
importable_data_dir = here("processing-and-analysis", "01-importable-data", "/")
analysis_data_dir = here("processing-and-analysis", "03-analysis-data", "/")
metadata_dir = here("original-data", "metadata", "/")
original_data_dir = here("original-data", "/")
```

### Copy rds file from original data directory

```{r}
#file.copy(paste0(original_data_dir, "master_2s_small_deidenitified.rds"),
#          paste0(importable_data_dir, "master_2s_small_deidenitified.rds"))

# CFC: Copy rds file from the `original_data_dir` to the `importable_data_dir`.
copy_from <- paste0(original_data_dir, "master_2s_small_deidentified.rds")
copy_to <- paste0(importable_data_dir, "master_2s_small_deidentified.rds")

file.copy(copy_from, copy_to)

# CFC: This line was used to determine if the original data file could be
# detected, since I kept running into errors. I finally fixed the
# issue...it was a spelling error in the file name.
# file.exists(copy_from)
```


### Import dataset

```{r}
# master = read_csv(paste0(original_data_dir, "master_2s_small_deidentified.csv"))

# CFC: IMport data set from `importable_data_dir`
master <- readRDS(paste0(importable_data_dir, "master_2s_small_deidenitified.rds"))

```

## Relationship 1: Scores for Exam 1, 2, and Final by:
* `course_fullid`
* `course_fullid` and `ver`
* `course_fullid` and demographic `eop_id`, `sex_id`, `fgn_id`, `urm_id` (all separate)

### Remove duplicate rows of student data:

```{r}
master_extras_removed = subset(master, select = -c(item_num, stud_ans, exam_key, corr, bloom_rating, complexity_rating_mean, complexity_rating_median, item_code, stem))

master_unique = unique(master_extras_removed, incomparables = FALSE)
```

### Faceted histograms of scores for exam 1, 2, and final by `course_fullid`:

```{r}
ggplot(master_unique, aes(x = exam1)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Exam 1 Score", y = "Number of Students", title = "Exam 1 Score Distributions by Course ID") +
  facet_wrap(~ course_fullid, nrow = 2)

ggplot(master_unique, aes(x = exam2)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Exam 2 Score", y = "Number of Students", title = "Exam 2 Score Distributions by Course ID") +
  facet_wrap(~ course_fullid, nrow = 2)

ggplot(master_unique, aes(x = finalexam)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Final Exam Score", y = "Number of Students", title = "Final Exam Score Distributions by Course ID") +
  facet_wrap(~ course_fullid, nrow = 2)
```

### Boxplots of Scores for Exam 1, 2, and final by `course_fullid`:

```{r}
ggplot(master_unique, aes(x = course_fullid, y = exam1)) +
  geom_boxplot() +
  labs(x = "Course ID", y = "Exam 1 Score", title = "Exam 1 Score Distributions by Course ID")

ggplot(master_unique, aes(x = course_fullid, y = exam2)) +
  geom_boxplot() +
  labs(x = "Course ID", y = "Exam 2 Score", title = "Exam 2 Score Distributions by Course ID")

ggplot(master_unique, aes(x = course_fullid, y = finalexam)) +
  geom_boxplot() +
  labs(x = "Course ID", y = "Final Exam Score", title = "Final Exam Score Distributions by Course ID")
```


### Faceted Histograms of Scores for Exam 1, 2, and final by `course_fullid` and `ver`:

```{r}
ggplot(master_unique, aes(x = exam1)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Exam 1 Score", y = "Number of Students", title = "Exam 1 Score Distributions by Course ID and Test Version") +
  facet_wrap(~ course_fullid + ver, nrow = 3)

ggplot(master_unique, aes(x = exam2)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Exam 2 Score", y = "Number of Students", title = "Exam 2 Score Distributions by Course ID and Test Version") +
  facet_wrap(~ course_fullid + ver, nrow = 3)

ggplot(master_unique, aes(x = finalexam)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Final Exam Score", y = "Number of Students", title = "Final Exam Score Distributions by Course ID and Test Version") +
  facet_wrap(~ course_fullid + ver, nrow = 3)
```

### Faceted Boxplots of Scores for Exam 1, 2, and final by `course_fullid` and `ver`:

```{r}
ggplot(master_unique, aes(x = course_fullid, y = exam1)) +
  geom_boxplot() +
  labs(x = "Course ID", y = "Exam 1 Score", title = "Exam 1 Score Distributions by Course ID") +
  facet_wrap(~ ver, nrow = 3)

ggplot(master_unique, aes(x = course_fullid, y = exam2)) +
  geom_boxplot() +
  labs(x = "Course ID", y = "Exam 2 Score", title = "Exam 2 Score Distributions by Course ID") +
  facet_wrap(~ ver, nrow = 3)

ggplot(master_unique, aes(x = course_fullid, y = finalexam)) +
  geom_boxplot() +
  labs(x = "Course ID", y = "Final Exam Score", title = "Final Exam Score Distributions by Course ID") +
  facet_wrap(~ ver, nrow = 3)
```


### Faceted Histograms of Scores for Exam 1, 2, and Final by `course_fullid` and demographic `eop_id`, `sex_id`, `fgn_id`, `urm_id`:

```{r}
ggplot(master_unique, aes(x = exam1)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Exam 1 Score", y = "Number of Students", title = "Exam 1 Score Distributions by Course ID and EOP ID") +
  facet_wrap(~ course_fullid + eop_id, nrow = 3)

ggplot(master_unique, aes(x = exam2)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Exam 2 Score", y = "Number of Students", title = "Exam 2 Score Distributions by Course ID and EOP ID") +
  facet_wrap(~ course_fullid + eop_id, nrow = 3)

ggplot(master_unique, aes(x = finalexam)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Final Exam Score", y = "Number of Students", title = "Final Exam Score Distributions by Course ID and EOP ID") +
  facet_wrap(~ course_fullid + eop_id, nrow = 3)

ggplot(master_unique, aes(x = exam1)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Exam 1 Score", y = "Number of Students", title = "Exam 1 Score Distributions by Course ID and Sex") +
  facet_wrap(~ course_fullid + sex_id, nrow = 3)

ggplot(master_unique, aes(x = exam2)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Exam 2 Score", y = "Number of Students", title = "Exam 2 Score Distributions by Course ID and Sex") +
  facet_wrap(~ course_fullid + sex_id, nrow = 3)

ggplot(master_unique, aes(x = finalexam)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Final Exam Score", y = "Number of Students", title = "Final Exam Score Distributions by Course ID and Sex") +
  facet_wrap(~ course_fullid + sex_id, nrow = 3)

ggplot(master_unique, aes(x = exam1)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Exam 1 Score", y = "Number of Students", title = "Exam 1 Score Distributions by Course ID and FGN ID") +
  facet_wrap(~ course_fullid + fgn_id, nrow = 3)

ggplot(master_unique, aes(x = exam2)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Exam 2 Score", y = "Number of Students", title = "Exam 2 Score Distributions by Course ID and FGN ID") +
  facet_wrap(~ course_fullid + fgn_id, nrow = 3)

ggplot(master_unique, aes(x = finalexam)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Final Exam Score", y = "Number of Students", title = "Final Exam Score Distributions by Course ID and FGN ID") +
  facet_wrap(~ course_fullid + fgn_id, nrow = 3)

ggplot(master_unique, aes(x = exam1)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Exam 1 Score", y = "Number of Students", title = "Exam 1 Score Distributions by Course ID and URM ID") +
  facet_wrap(~ course_fullid + urm_id, nrow = 3)

ggplot(master_unique, aes(x = exam2)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Exam 2 Score", y = "Number of Students", title = "Exam 2 Score Distributions by Course ID and URM ID") +
  facet_wrap(~ course_fullid + urm_id, nrow = 3)

ggplot(master_unique, aes(x = finalexam)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Final Exam Score", y = "Number of Students", title = "Final Exam Score Distributions by Course ID and URM ID") +
  facet_wrap(~ course_fullid + urm_id, nrow = 3)
```

### Faceted Boxplots of Scores for Exam 1, 2, and Final by `course_fullid` and demographic `eop_id`, `sex_id`, `fgn_id`, `urm_id`:

```{r}
ggplot(master_unique, aes(x = course_fullid, y = exam1)) +
  geom_boxplot() +
  labs(x = "Course ID", y = "Exam 1 Score", title = "Exam 1 Score Distributions by Course ID and EOP ID") +
  facet_wrap(~ eop_id, nrow = 3)

ggplot(master_unique, aes(x = course_fullid, y = exam2)) +
  geom_boxplot() +
  labs(x = "Course ID", y = "Exam 2 Score", title = "Exam 2 Score Distributions by Course ID and EOP ID") +
  facet_wrap(~ eop_id, nrow = 3)

ggplot(master_unique, aes(x = course_fullid, y = finalexam)) +
  geom_boxplot() +
  labs(x = "Course ID", y = "Final Exam Score", title = "Final Exam Score Distributions by Course ID and EOP ID") +
  facet_wrap(~ eop_id, nrow = 3)

ggplot(master_unique, aes(x = course_fullid, y = exam1)) +
  geom_boxplot() +
  labs(x = "Course ID", y = "Exam 1 Score", title = "Exam 1 Score Distributions by Course ID and Sex") +
  facet_wrap(~ sex_id, nrow = 3)

ggplot(master_unique, aes(x = course_fullid, y = exam2)) +
  geom_boxplot() +
  labs(x = "Course ID", y = "Exam 2 Score", title = "Exam 2 Score Distributions by Course ID and Sex") +
  facet_wrap(~ sex_id, nrow = 3)

ggplot(master_unique, aes(x = course_fullid, y = finalexam)) +
  geom_boxplot() +
  labs(x = "Course ID", y = "Final Exam Score", title = "Final Exam Score Distributions by Course ID and Sex") +
  facet_wrap(~ sex_id, nrow = 3)

ggplot(master_unique, aes(x = course_fullid, y = exam1)) +
  geom_boxplot() +
  labs(x = "Course ID", y = "Exam 1 Score", title = "Exam 1 Score Distributions by Course ID and FGN ID") +
  facet_wrap(~ fgn_id, nrow = 3)

ggplot(master_unique, aes(x = course_fullid, y = exam2)) +
  geom_boxplot() +
  labs(x = "Course ID", y = "Exam 2 Score", title = "Exam 2 Score Distributions by Course ID and FGN ID") +
  facet_wrap(~ fgn_id, nrow = 3)

ggplot(master_unique, aes(x = course_fullid, y = finalexam)) +
  geom_boxplot() +
  labs(x = "Course ID", y = "Final Exam Score", title = "Final Exam Score Distributions by Course ID and FGN ID") +
  facet_wrap(~ fgn_id, nrow = 3)

ggplot(master_unique, aes(x = course_fullid, y = exam1)) +
  geom_boxplot() +
  labs(x = "Course ID", y = "Exam 1 Score", title = "Exam 1 Score Distributions by Course ID and URM ID") +
  facet_wrap(~ urm_id, nrow = 3)

ggplot(master_unique, aes(x = course_fullid, y = exam2)) +
  geom_boxplot() +
  labs(x = "Course ID", y = "Exam 2 Score", title = "Exam 2 Score Distributions by Course ID and URM ID") +
  facet_wrap(~ urm_id, nrow = 3)

ggplot(master_unique, aes(x = course_fullid, y = finalexam)) +
  geom_boxplot() +
  labs(x = "Course ID", y = "Final Exam Score", title = "Final Exam Score Distributions by Course ID and URM ID") +
  facet_wrap(~ urm_id, nrow = 3)
```

## Relationship 2: Scores for each of the A17 individual quiz scores (`q1_ind`) by:
* `ver`
* `eop_id`, `sex_id`, `fgn_id`, `urm_id`

### Condense dataset to only include Autumn 2017 data:

```{r}
fall_2017 = master %>% 
  filter(course_fullid == "CHEM_142_A_2017_4" | course_fullid == "CHEM_142_B_2017_4")
fall_2017 = na.omit(fall_2017)
```

### Faceted histograms of `q1_ind` by `ver`:

```{r}
ggplot(fall_2017, aes(x = q1_ind)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Individual Quiz Score", y = "Number of Students", title = "Individual Quiz Score Distributions by Version") +
  facet_wrap(~ ver, nrow = 2)
```

### Boxplots of `q1_ind` by `ver`:

```{r}
ggplot(fall_2017, aes(x = ver, y = q1_ind)) +
  geom_boxplot() +
  labs(x = "Version", y = "Individual Quiz Scores", title = "Boxplots of Individual Quiz Scores by Version")
```

### Faceted Histograms of `q1_ind` by `eop_id`, `sex_id`, `fgn_id`, and `urm_id`: 

```{r}
ggplot(fall_2017, aes(x = q1_ind)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Individual Quiz Score", y = "number of students", title = "Distributions of Individual Quiz Scores by EOP ID") +
  facet_wrap(~ eop_id, nrow = 1)

ggplot(fall_2017, aes(x = q1_ind)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Individual Quiz Score", y = "number of students", title = "Distributions of Individual Quiz Scores by Sex") +
  facet_wrap(~ sex_id, nrow = 1)

ggplot(fall_2017, aes(x = q1_ind)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Individual Quiz Score", y = "number of students", title = "Distributions of Individual Quiz Scores by FGN ID") +
  facet_wrap(~ fgn_id, nrow = 1)

ggplot(fall_2017, aes(x = q1_ind)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Individual Quiz Score", y = "number of students", title = "Distributions of Individual Quiz Scores by URM ID") +
  facet_wrap(~ urm_id, nrow = 1)
```

### Boxplots of `q1_ind` by `eop_id`, `sex_id`, `fgn_id`, and `urm_id`: 

```{r}
ggplot(fall_2017, aes(x = eop_id, y = q1_ind)) +
  geom_boxplot() +
  labs(x = "EOP ID", y = "Individual Quiz Scores", title = "Distributions of Individual Quiz Scores by EOP ID")

ggplot(fall_2017, aes(x = sex_id, y = q1_ind)) +
  geom_boxplot() +
  labs(x = "Sex", y = "Individual Quiz Scores", title = "Distributions of Individual Quiz Scores by Sex")

ggplot(fall_2017, aes(x = fgn_id, y = q1_ind)) +
  geom_boxplot() +
  labs(x = "FGN ID", y = "Individual Quiz Scores", title = "Distributions of Individual Quiz Scores by FGN ID")

ggplot(fall_2017, aes(x = urm_id, y = q1_ind)) +
  geom_boxplot() +
  labs(x = "URM ID", y = "Individual Quiz Scores", title = "Distributions of Individual Quiz Scores by URM ID")
```

## Relationship 3: Scores for each of the A17 group quiz scores (`q1_grp`) by:
* `ver`
* `eop_id`, `sex_id`, `fgn_id`, `urm_id`

### Faceted histograms of `q1_grp` by `ver`:

```{r}
ggplot(fall_2017, aes(x = q1_grp)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Group Quiz Score", y = "Number of Students", title = "Group Quiz Score Distributions by Version") +
  facet_wrap(~ ver, nrow = 2)
```

### Boxplots of `q1_grp` by `ver`:

```{r}
ggplot(fall_2017, aes(x = ver, y = q1_grp)) +
  geom_boxplot() +
  labs(x = "Version", y = "Group Quiz Scores", title = "Boxplots of Group Quiz Scores by Version")
```

### Faceted Histograms of `q1_grp` by `eop_id`, `sex_id`, `fgn_id`, and `urm_id`: 

```{r}
ggplot(fall_2017, aes(x = q1_grp)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Group Quiz Score", y = "number of students", title = "Distributions of Group Quiz Scores by EOP ID") +
  facet_wrap(~ eop_id, nrow = 1)

ggplot(fall_2017, aes(x = q1_grp)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Group Quiz Score", y = "number of students", title = "Distributions of Group Quiz Scores by Sex") +
  facet_wrap(~ sex_id, nrow = 1)

ggplot(fall_2017, aes(x = q1_grp)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Group Quiz Score", y = "number of students", title = "Distributions of Group Quiz Scores by FGN ID") +
  facet_wrap(~ fgn_id, nrow = 1)

ggplot(fall_2017, aes(x = q1_grp)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Group Quiz Score", y = "number of students", title = "Distributions of Group Quiz Scores by URM ID") +
  facet_wrap(~ urm_id, nrow = 1)
```

### Boxplots of `q1_grp` by `eop_id`, `sex_id`, `fgn_id`, and `urm_id`: 

```{r}
ggplot(fall_2017, aes(x = eop_id, y = q1_grp)) +
  geom_boxplot() +
  labs(x = "EOP ID", y = "Group Quiz Scores", title = "Distributions of Group Quiz Scores by EOP ID")

ggplot(fall_2017, aes(x = sex_id, y = q1_grp)) +
  geom_boxplot() +
  labs(x = "Sex", y = "Group Quiz Scores", title = "Distributions of Group Quiz Scores by Sex")

ggplot(fall_2017, aes(x = fgn_id, y = q1_grp)) +
  geom_boxplot() +
  labs(x = "FGN ID", y = "Group Quiz Scores", title = "Distributions of Group Quiz Scores by FGN ID")

ggplot(fall_2017, aes(x = urm_id, y = q1_grp)) +
  geom_boxplot() +
  labs(x = "URM ID", y = "Group Quiz Scores", title = "Distributions of Group Quiz Scores by URM ID")
```

## Relationship 4: Scores for each of the A17 total quiz scores (`q1_total`) by:
* `ver`
* `eop_id`, `sex_id`, `fgn_id`, `urm_id`

### Faceted histograms of `q1_total` by `ver`:

```{r}
ggplot(fall_2017, aes(x = q1_total)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Total Quiz Score", y = "Number of Students", title = "Total Quiz Score Distributions by Version") +
  facet_wrap(~ ver, nrow = 2)
```

### Boxplots of `q1_total` by `ver`:

```{r}
ggplot(fall_2017, aes(x = ver, y = q1_total)) +
  geom_boxplot() +
  labs(x = "Version", y = "Total Quiz Scores", title = "Boxplots of Total Quiz Scores by Version")
```

### Faceted Histograms of `q1_total` by `eop_id`, `sex_id`, `fgn_id`, and `urm_id`: 

```{r}
ggplot(fall_2017, aes(x = q1_total)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Total Quiz Score", y = "number of students", title = "Distributions of Total Quiz Scores by EOP ID") +
  facet_wrap(~ eop_id, nrow = 1)

ggplot(fall_2017, aes(x = q1_total)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Total Quiz Score", y = "number of students", title = "Distributions of Total Quiz Scores by Sex") +
  facet_wrap(~ sex_id, nrow = 1)

ggplot(fall_2017, aes(x = q1_total)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Total Quiz Score", y = "number of students", title = "Distributions of Total Quiz Scores by FGN ID") +
  facet_wrap(~ fgn_id, nrow = 1)

ggplot(fall_2017, aes(x = q1_total)) +
  geom_histogram(binwidth = 5, color = "white") +
  labs(x = "Total Quiz Score", y = "number of students", title = "Distributions of Total Quiz Scores by URM ID") +
  facet_wrap(~ urm_id, nrow = 1)
```

### Boxplots of `q1_total` by `eop_id`, `sex_id`, `fgn_id`, and `urm_id`: 

```{r}
ggplot(fall_2017, aes(x = eop_id, y = q1_total)) +
  geom_boxplot() +
  labs(x = "EOP ID", y = "Total Quiz Scores", title = "Distributions of Total Quiz Scores by EOP ID")

ggplot(fall_2017, aes(x = sex_id, y = q1_total)) +
  geom_boxplot() +
  labs(x = "Sex", y = "Total Quiz Scores", title = "Distributions of Total Quiz Scores by Sex")

ggplot(fall_2017, aes(x = fgn_id, y = q1_total)) +
  geom_boxplot() +
  labs(x = "FGN ID", y = "Total Quiz Scores", title = "Distributions of Total Quiz Scores by FGN ID")

ggplot(fall_2017, aes(x = urm_id, y = q1_total)) +
  geom_boxplot() +
  labs(x = "URM ID", y = "Total Quiz Scores", title = "Distributions of Total Quiz Scores by URM ID")
```

## Summary Statistics:

### Exam Scores by `course_fullid`:
```{r}
fall_2016_A = filter(master_unique, course_fullid == "CHEM_142_A_2016_4")
fall_2016_B = filter(master_unique, course_fullid == "CHEM_142_B_2016_4")
fall_2017_A = filter(master_unique, course_fullid == "CHEM_142_A_2017_4")
fall_2017_B = filter(master_unique, course_fullid == "CHEM_142_B_2017_4")

summary_exam1_fall_2016_A = fall_2016_A %>% 
  summarize(mean = mean(exam1), std_dev = sd(exam1), min = min(exam1), max = max(exam1))
summary_exam1_fall_2016_A
summary_exam2_fall_2016_A = fall_2016_A %>% 
  summarize(mean = mean(exam2), std_dev = sd(exam2), min = min(exam2), max = max(exam2))
summary_exam2_fall_2016_A
summary_final_fall_2016_A = fall_2016_A %>% 
  summarize(mean = mean(finalexam), std_dev = sd(finalexam), min = min(finalexam), max = max(finalexam))
summary_final_fall_2016_A

summary_exam1_fall_2016_B = fall_2016_B %>% 
  summarize(mean = mean(exam1), std_dev = sd(exam1), min = min(exam1), max = max(exam1))
summary_exam1_fall_2016_B
summary_exam2_fall_2016_B = fall_2016_B %>% 
  summarize(mean = mean(exam2), std_dev = sd(exam2), min = min(exam2), max = max(exam2))
summary_exam2_fall_2016_B
summary_final_fall_2016_B = fall_2016_B %>% 
  summarize(mean = mean(finalexam), std_dev = sd(finalexam), min = min(finalexam), max = max(finalexam))
summary_final_fall_2016_B

summary_exam1_fall_2017_A = fall_2017_A %>% 
  summarize(mean = mean(exam1), std_dev = sd(exam1), min = min(exam1), max = max(exam1))
summary_exam1_fall_2017_A
summary_exam2_fall_2017_A = fall_2017_A %>% 
  summarize(mean = mean(exam2), std_dev = sd(exam2), min = min(exam2), max = max(exam2))
summary_exam2_fall_2017_A
summary_final_fall_2017_A = fall_2017_A %>% 
  summarize(mean = mean(finalexam), std_dev = sd(finalexam), min = min(finalexam), max = max(finalexam))
summary_final_fall_2017_A

summary_exam1_fall_2017_B = fall_2017_B %>% 
  summarize(mean = mean(exam1), std_dev = sd(exam1), min = min(exam1), max = max(exam1))
summary_exam1_fall_2017_B
summary_exam2_fall_2017_B = fall_2017_B %>% 
  summarize(mean = mean(exam2), std_dev = sd(exam2), min = min(exam2), max = max(exam2))
summary_exam2_fall_2017_B
summary_final_fall_2017_B = fall_2017_B %>% 
  summarize(mean = mean(finalexam), std_dev = sd(finalexam), min = min(finalexam), max = max(finalexam))
summary_final_fall_2017_B
```

### Exam Scores by `course_fullid` and `eop_id`:
```{r}
summary_exam1_fall_2016_A_eop = fall_2016_A %>%
  filter(eop_id == "EOP") %>% 
  summarize(mean = mean(exam1), st_dev = sd(exam1), min = min(exam1), max = max(exam1))
summary_exam1_fall_2016_A_eop

summary_exam1_fall_2016_A_non_eop = fall_2016_A %>%
  filter(eop_id == "non-EOP") %>% 
  summarize(mean = mean(exam1), st_dev = sd(exam1), min = min(exam1), max = max(exam1))
summary_exam1_fall_2016_A_non_eop
```

### Exam Scores by `course_fullid` and `sex_id`:
```{r}
summary_exam1_fall_2016_A_male = fall_2016_A %>%
  filter(sex_id == "Male") %>% 
  summarize(mean = mean(exam1), st_dev = sd(exam1), min = min(exam1), max = max(exam1))
summary_exam1_fall_2016_A_male

summary_exam1_fall_2016_A_female = fall_2016_A %>%
  filter(eop_id == "Female") %>% 
  summarize(mean = mean(exam1), st_dev = sd(exam1), min = min(exam1), max = max(exam1))
summary_exam1_fall_2016_A_female
```

### Exam Scores by `course_fullid` and `fgn_id`:
```{r}
summary_exam1_fall_2016_A_fgn = fall_2016_A %>%
  filter(fgn_id == "FGN") %>% 
  summarize(mean = mean(exam1), st_dev = sd(exam1), min = min(exam1), max = max(exam1))
summary_exam1_fall_2016_A_fgn

summary_exam1_fall_2016_A_non_fgn = fall_2016_A %>%
  filter(fgn_id == "non-FGN") %>% 
  summarize(mean = mean(exam1), st_dev = sd(exam1), min = min(exam1), max = max(exam1))
summary_exam1_fall_2016_A_non_fgn
```

### Exam scores by `course_fullid` and `urm_id`:
```{r}
summary_exam1_fall_2016_A_urm = fall_2016_A %>%
  filter(urm_id == "URM") %>% 
  summarize(mean = mean(exam1), st_dev = sd(exam1), min = min(exam1), max = max(exam1))
summary_exam1_fall_2016_A_urm

summary_exam1_fall_2016_A_non_urm = fall_2016_A %>%
  filter(urm_id == "non-URM") %>% 
  summarize(mean = mean(exam1), st_dev = sd(exam1), min = min(exam1), max = max(exam1))
summary_exam1_fall_2016_A_non_urm
```

## Notes:
* Need to figure out how to remove the NAs
* Still need to include summary stats for the quiz data
* The dataset is still being uploaded through my local device (I forgot how to upload properly through the file stream)