---
title: "Playing with `effects` package"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(effects)
```

# Cowles and Davis's Data on Volunteering
## Description
The Cowles data frame has 1421 rows and 4 columns. These data come from a study of the personality determinants of volunteering for psychological research.

## Format
This data frame contains the following columns:

* **neuroticism**: scale from Eysenck personality inventory
* **extraversion**: scale from Eysenck personality inventory
* **sex**: a factor with levels: female; male
* **volunteer**: volunteering, a factor with levels: no; yes

## Source
Cowles, M. and C. Davis (1987) The subject matter of psychology: Volunteers. 
_British Journal of Social Psychology_ 26, 97–102.

```{r}
mod.cowles <- glm(volunteer ~ sex + neuroticism*extraversion,
                  data = Cowles, family = binomial)

eff.cowles <- allEffects(mod.cowles, 
                         xlevels = list(extraversion = seq(0, 24, 6)),
                         fixed.predictors = list(given.values = c(sexmale = 0.5)))

eff.cowles_2 <- allEffects(mod.cowles, 
                         xlevels = list(extraversion = seq(0, 24, 6)),
                         fixed.predictors = list(given.values = c(sexmale = 1.0)))
```

```{r}
summary(mod.cowles)
summary.eff(eff.cowles)
```


```{r}

eff.cowles
eff.cowles_2
as.data.frame(eff.cowles[[1]])
as.data.frame(eff.cowles[[2]])
as.data.frame(eff.cowles_2[[1]])
as.data.frame(eff.cowles_2[[2]])
```

```{r}
# the following are equivalent:
eff.ne <- effect("neuroticism*extraversion", mod.cowles)
Eff.ne <- Effect(c("neuroticism", "extraversion"), mod.cowles)
all.equal(eff.ne$fit, Eff.ne$fit)

plot(eff.cowles, 'sex', axes=list(y=list(lab="Prob(Volunteer)")))

plot(eff.cowles, 'neuroticism:extraversion',
     axes=list(y=list(lab="Prob(Volunteer)",
        ticks=list(at=c(.1,.25,.5,.75,.9)))))

plot(Effect(c("neuroticism", "extraversion"), mod.cowles,
            se=list(type="Scheffe"),
            xlevels=list(extraversion=seq(0, 24, 6)),
            fixed.predictors=list(given.values=c(sexmale=0.5))),
     axes=list(y=list(lab="Prob(Volunteer)",
        ticks=list(at=c(.1,.25,.5,.75,.9)))))

plot(eff.cowles, 'neuroticism:extraversion', lines=list(multiline=TRUE),
     axes=list(y=list(lab="Prob(Volunteer)")))

plot(effect('sex:neuroticism:extraversion', mod.cowles,
            xlevels=list(extraversion=seq(0, 24, 6))),
     lines=list(multiline=TRUE))

```

