---
title: "data_2s_summary"
author: "Ganling"
date: "4/19/2021"
output: 
  html_document:
    keep_md: yes
---

#Load the packages
```{r}
library(here)
library(tidyverse)
library(skimr)
library(moderndive)
library(ggplot2)
library(dplyr)
library(readr)
```
#read the files using here package
```{r}
proj_dir <- here()
importable_data_dir <- here("processing-and-analysis", "importable-data", "/")
analysis_data_dir <- here("processing-and-analysis", "analysis-data", "/")
metadata_dir <- here("original-data", "metadata", "/")

`master_2s_small_deidentified (DO NOT TOUCH)` <- readRDS("G:/Shared drives/CEG R Coding Practice/Ganling/two-stage-2/processing-and-analysis/01-importable-data/dataset/master_2s_small_deidentified (DO NOT TOUCH).rds")

data_2s <- `master_2s_small_deidentified (DO NOT TOUCH)`
```
###Before starting to make historgrams and boxplots, I would like to remove all the duplicating rows and make a new data frame
```{r}
data_2s_extras_removed = subset(data_2s, select = -c(item_num, stud_ans, exam_key, corr, bloom_rating, complexity_rating_mean, complexity_rating_median, item_code, stem))

data_2s_unique = unique(data_2s_extras_removed, incomparables = FALSE)
View(data_2s_unique)

data_2s_exam_score <- data_2s_unique %>% select(exam1, exam2, finalexam, course_fullid, ver, sex_id, urm_id, eop_id, fgn_id)

data_2s_exam_score_true <- na.omit(data_2s_exam_score)
```
#Now I would like to graph the histogram and boxplots for exam scores and course_id
```{r}
#histogram for coruse_id
ggplot(data = data_2s_exam_score_true, mapping = aes(x= exam1)) + geom_histogram(color = "white") + labs(x = "Exam 1 score", y= "number of students", title = ("Exam 1 score Distributions by Course ID")) + facet_wrap(~ course_fullid, nrow = 1 )

ggplot(data = data_2s_exam_score_true, mapping = aes(x= exam2)) + geom_histogram(color = "white") + labs(x = "Exam 2 score", y= "number of students", title = ("Exam 2 score Distributions by Course ID")) + facet_wrap(~ course_fullid, nrow = 1 )

ggplot(data = data_2s_exam_score_true, mapping = aes(x= finalexam)) + geom_histogram(color = "white") + labs(x = "Final exam score", y= "number of students", title = ("Final exam score Distributions by Course ID")) + facet_wrap(~ course_fullid, nrow = 1 )

#boxplot for course_id
ggplot(data = data_2s_exam_score_true, mapping = aes(x= course_fullid, y= exam1)) + geom_boxplot() + labs(x = "Exam 1 score", y= "number of students", title = ("Exam 1 score Distributions by Course ID")) + facet_wrap(~ course_fullid, nrow = 1 )

ggplot(data = data_2s_exam_score_true, mapping = aes(x= course_fullid, y= exam2)) + geom_boxplot() + labs(x = "Exam 2 score", y= "number of students", title = ("Exam 2 score Distributions by Course ID")) + facet_wrap(~ course_fullid, nrow = 1 )

ggplot(data = data_2s_exam_score_true, mapping = aes(x= course_fullid, y= finalexam)) + geom_boxplot() + labs(x = "Final exam score", y= "number of students", title = ("Final exam score Distributions by Course ID")) + facet_wrap(~ course_fullid, nrow = 1 )
```
# next, we want to look at exam score, course id and version 
```{r}
#histogram for coruse_id
ggplot(data = data_2s_exam_score_true, mapping = aes(x= exam1)) + geom_histogram(color = "white") + labs(x = "Exam 1 score", y= "number of students", title = ("Exam 1 score Distributions by Course ID and version")) + facet_wrap(~ course_fullid + ver, nrow = 3 )

ggplot(data = data_2s_exam_score_true, mapping = aes(x= exam2)) + geom_histogram(color = "white") + labs(x = "Exam 2 score", y= "number of students", title = ("Exam 2 score Distributions by Course ID and version")) + facet_wrap(~ course_fullid + ver, nrow = 3 )

ggplot(data = data_2s_exam_score_true, mapping = aes(x= finalexam)) + geom_histogram(color = "white") + labs(x = "Final exam score", y= "number of students", title = ("Final exam score Distributions by Course ID and version")) + facet_wrap(~ course_fullid + ver, nrow = 3 )

#boxplot for course_id
ggplot(data = data_2s_exam_score_true, mapping = aes(x= course_fullid, y= exam1)) + geom_boxplot() + labs(x = "Exam 1 score", y= "number of students", title = ("Exam 1 score Distributions by Course ID and version")) + facet_wrap(~ course_fullid + ver, nrow = 3 )

ggplot(data = data_2s_exam_score_true, mapping = aes(x= course_fullid, y= exam2)) + geom_boxplot() + labs(x = "Exam 2 score", y= "number of students", title = ("Exam 2 score Distributions by Course ID and version")) + facet_wrap(~ course_fullid + ver, nrow = 3 )

ggplot(data = data_2s_exam_score_true, mapping = aes(x= course_fullid, y= finalexam)) + geom_boxplot() + labs(x = "Final exam score", y= "number of students", title = ("Final exam score Distributions by Course ID and version")) + facet_wrap(~ course_fullid + ver, nrow = 3 )
```
#course id and eop status
```{r}
#histogram for coruse_id
ggplot(data = data_2s_exam_score_true, mapping = aes(x= exam1)) + geom_histogram(color = "white") + labs(x = "Exam 1 score", y= "number of students", title = ("Exam 1 score Distributions by Course ID and eop")) + facet_wrap(~ course_fullid + eop_id, nrow = 3 )

ggplot(data = data_2s_exam_score_true, mapping = aes(x= exam2)) + geom_histogram(color = "white") + labs(x = "Exam 2 score", y= "number of students", title = ("Exam 2 score Distributions by Course ID and eop")) + facet_wrap(~ course_fullid + eop_id, nrow = 3 )

ggplot(data = data_2s_exam_score_true, mapping = aes(x= finalexam)) + geom_histogram(color = "white") + labs(x = "Final exam score", y= "number of students", title = ("Final exam score Distributions by Course ID and eop")) + facet_wrap(~ course_fullid + eop_id, nrow = 3 )

#boxplot for course_id
ggplot(data = data_2s_exam_score_true, mapping = aes(y= exam1)) + geom_boxplot() + labs(x = "Exam 1 score", y= "number of students", title = ("Exam 1 score Distributions by Course ID and eop")) + facet_wrap(~ course_fullid + eop_id, nrow = 3 )

ggplot(data = data_2s_exam_score_true, mapping = aes(y= exam2)) + geom_boxplot() + labs(x = "Exam 2 score", y= "number of students", title = ("Exam 2 score Distributions by Course ID and eop")) + facet_wrap(~ course_fullid + eop_id, nrow = 3 )

ggplot(data = data_2s_exam_score_true, mapping = aes(y= finalexam)) + geom_boxplot() + labs(x = "Final exam score", y= "number of students", title = ("Final exam score Distributions by Course ID and eop")) + facet_wrap(~ course_fullid + eop_id, nrow = 3 )
```
#course id and sex id
```{r}
#histogram for coruse_id
ggplot(data = data_2s_exam_score_true, mapping = aes(x= exam1)) + geom_histogram(color = "white") + labs(x = "Exam 1 score", y= "number of students", title = ("Exam 1 score Distributions by Course ID and sex")) + facet_wrap(~ course_fullid + sex_id, nrow = 3 )

ggplot(data = data_2s_exam_score_true, mapping = aes(x= exam2)) + geom_histogram(color = "white") + labs(x = "Exam 2 score", y= "number of students", title = ("Exam 2 score Distributions by Course ID and sex")) + facet_wrap(~ course_fullid + sex_id, nrow = 3 )

ggplot(data = data_2s_exam_score_true, mapping = aes(x= finalexam)) + geom_histogram(color = "white") + labs(x = "Final exam score", y= "number of students", title = ("Final exam score Distributions by Course ID and sex")) + facet_wrap(~ course_fullid + sex_id, nrow = 3 )

#boxplot for course_id
ggplot(data = data_2s_exam_score_true, mapping = aes(y= exam1)) + geom_boxplot() + labs(x = "Exam 1 score", y= "number of students", title = ("Exam 1 score Distributions by Course ID and sex")) + facet_wrap(~ course_fullid + sex_id, nrow = 3 )

ggplot(data = data_2s_exam_score_true, mapping = aes(y= exam2)) + geom_boxplot() + labs(x = "Exam 2 score", y= "number of students", title = ("Exam 2 score Distributions by Course ID and sex")) + facet_wrap(~ course_fullid + sex_id, nrow = 3 )

ggplot(data = data_2s_exam_score_true, mapping = aes(y= finalexam)) + geom_boxplot() + labs(x = "Final exam score", y= "number of students", title = ("Final exam score Distributions by Course ID and sex")) + facet_wrap(~ course_fullid + sex_id, nrow = 3 )
```
#course id and fgn status
```{r}
#histogram for coruse_id
ggplot(data = data_2s_exam_score_true, mapping = aes(x= exam1)) + geom_histogram(color = "white") + labs(x = "Exam 1 score", y= "number of students", title = ("Exam 1 score Distributions by Course ID and fgn")) + facet_wrap(~ course_fullid + fgn_id, nrow = 3 )

ggplot(data = data_2s_exam_score_true, mapping = aes(x= exam2)) + geom_histogram(color = "white") + labs(x = "Exam 2 score", y= "number of students", title = ("Exam 2 score Distributions by Course ID and fgn")) + facet_wrap(~ course_fullid + fgn_id, nrow = 3 )

ggplot(data = data_2s_exam_score_true, mapping = aes(x= finalexam)) + geom_histogram(color = "white") + labs(x = "Final exam score", y= "number of students", title = ("Final exam score Distributions by Course ID and fgn")) + facet_wrap(~ course_fullid + fgn_id, nrow = 3 )

#boxplot for course_id
ggplot(data = data_2s_exam_score_true, mapping = aes(y= exam1)) + geom_boxplot() + labs(x = "Exam 1 score", y= "number of students", title = ("Exam 1 score Distributions by Course ID and fgn")) + facet_wrap(~ course_fullid + fgn_id, nrow = 3 )

ggplot(data = data_2s_exam_score_true, mapping = aes(y= exam2)) + geom_boxplot() + labs(x = "Exam 2 score", y= "number of students", title = ("Exam 2 score Distributions by Course ID and fgn")) + facet_wrap(~ course_fullid + fgn_id, nrow = 3 )

ggplot(data = data_2s_exam_score_true, mapping = aes(y= finalexam)) + geom_boxplot() + labs(x = "Final exam score", y= "number of students", title = ("Final exam score Distributions by Course ID and fgn")) + facet_wrap(~ course_fullid + fgn_id, nrow = 3 )
```
#urm status and course id
```{r}
#histogram for coruse_id
ggplot(data = data_2s_exam_score_true, mapping = aes(x= exam1)) + geom_histogram(color = "white") + labs(x = "Exam 1 score", y= "number of students", title = ("Exam 1 score Distributions by Course ID and urm")) + facet_wrap(~ course_fullid + urm_id, nrow = 3 )

ggplot(data = data_2s_exam_score_true, mapping = aes(x= exam2)) + geom_histogram(color = "white") + labs(x = "Exam 2 score", y= "number of students", title = ("Exam 2 score Distributions by Course ID and urm")) + facet_wrap(~ course_fullid + urm_id, nrow = 3 )

ggplot(data = data_2s_exam_score_true, mapping = aes(x= finalexam)) + geom_histogram(color = "white") + labs(x = "Final exam score", y= "number of students", title = ("Final exam score Distributions by Course ID and urm")) + facet_wrap(~ course_fullid + urm_id, nrow = 3 )

#boxplot for course_id
ggplot(data = data_2s_exam_score_true, mapping = aes(y= exam1)) + geom_boxplot() + labs(x = "Exam 1 score", y= "number of students", title = ("Exam 1 score Distributions by Course ID and urm")) + facet_wrap(~ course_fullid + urm_id, nrow = 3 )

ggplot(data = data_2s_exam_score_true, mapping = aes(y= exam2)) + geom_boxplot() + labs(x = "Exam 2 score", y= "number of students", title = ("Exam 2 score Distributions by Course ID and urm")) + facet_wrap(~ course_fullid + urm_id, nrow = 3 )

ggplot(data = data_2s_exam_score_true, mapping = aes(y= finalexam)) + geom_boxplot() + labs(x = "Final exam score", y= "number of students", title = ("Final exam score Distributions by Course ID and urm")) + facet_wrap(~ course_fullid + urm_id, nrow = 3 )
```
#Now I want to look at A17 individual quiz score by ver, eop_id, sex_id, fgn_id and urm_id
```{r}
#filter out the right data set
data_2s_exam_score_true_A17 <- data_2s_unique %>% filter(course_fullid %in% c("CHEM_142_A_2017_4")) %>% select(q1_ind, q1_grp, q1_total, course_fullid, ver, eop_id, sex_id, fgn_id, urm_id)

ggplot(data = data_2s_exam_score_true_A17, mapping = aes(x= q1_ind)) + geom_histogram(color = "white") + labs(x = "Quiz 1 Individual score", y= "number of students", title = ("Quiz 1 individual score Distributions by version during A17")) + facet_wrap(~ ver, nrow = 3 )

#by eop
ggplot(data = data_2s_exam_score_true_A17, mapping = aes(x= q1_ind)) + geom_histogram(color = "white") + labs(x = "Quiz 1 Individual score", y= "number of students", title = ("Quiz 1 individual score Distributions by eop during A17")) + facet_wrap(~ eop_id, nrow = 3 )

#by sex
ggplot(data = data_2s_exam_score_true_A17, mapping = aes(x= q1_ind)) + geom_histogram(color = "white") + labs(x = "Quiz 1 Individual score", y= "number of students", title = ("Quiz 1 individual score Distributions by sex during A17")) + facet_wrap(~ sex_id, nrow = 3 )

#by fgn
ggplot(data = data_2s_exam_score_true_A17, mapping = aes(x= q1_ind)) + geom_histogram(color = "white") + labs(x = "Quiz 1 Individual score", y= "number of students", title = ("Quiz 1 individual score Distributions by first generation status during A17")) + facet_wrap(~ fgn_id, nrow = 3 )

#by urm
ggplot(data = data_2s_exam_score_true_A17, mapping = aes(x= q1_ind)) + geom_histogram(color = "white") + labs(x = "Quiz 1 Individual score", y= "number of students", title = ("Quiz 1 individual score Distributions by representation during A17")) + facet_wrap(~ urm_id, nrow = 3 )

##boxplots
ggplot(data = data_2s_exam_score_true_A17, mapping = aes(y= q1_ind)) + geom_boxplot() + labs(x = "Quiz 1 Individual score", y= "number of students", title = ("Quiz 1 individual score Distributions by version during A17")) + facet_wrap(~ ver, nrow = 3 )

#by eop
ggplot(data = data_2s_exam_score_true_A17, mapping = aes(y= q1_ind)) + geom_boxplot() + labs(x = "Quiz 1 Individual score", y= "number of students", title = ("Quiz 1 individual score Distributions by eop during A17")) + facet_wrap(~ eop_id, nrow = 3 )

#by sex
ggplot(data = data_2s_exam_score_true_A17, mapping = aes(y= q1_ind)) + geom_boxplot() + labs(x = "Quiz 1 Individual score", y= "number of students", title = ("Quiz 1 individual score Distributions by sex during A17")) + facet_wrap(~ sex_id, nrow = 3 )

#by fgn
ggplot(data = data_2s_exam_score_true_A17, mapping = aes(y= q1_ind)) + geom_boxplot() + labs(x = "Quiz 1 Individual score", y= "number of students", title = ("Quiz 1 individual score Distributions by first generation status during A17")) + facet_wrap(~ fgn_id, nrow = 3 )

#by urm
ggplot(data = data_2s_exam_score_true_A17, mapping = aes(y= q1_ind)) + geom_boxplot() + labs(x = "Quiz 1 Individual score", y= "number of students", title = ("Quiz 1 individual score Distributions by representation during A17")) + facet_wrap(~ urm_id, nrow = 3 )

```


#Now I want to look at A17 individual quiz score by ver, eop_id, sex_id, fgn_id and urm_id
```{r}
ggplot(data = data_2s_exam_score_true_A17, mapping = aes(x= q1_grp)) + geom_histogram(color = "white") + labs(x = "Quiz 1 group score", y= "number of students", title = ("Quiz 1 group score Distributions by version during A17")) + facet_wrap(~ ver, nrow = 3 )

#by eop
ggplot(data = data_2s_exam_score_true_A17, mapping = aes(x= q1_grp)) + geom_histogram(color = "white") + labs(x = "Quiz 1 group score", y= "number of students", title = ("Quiz 1 group score Distributions by eop during A17")) + facet_wrap(~ eop_id, nrow = 3 )

#by sex
ggplot(data = data_2s_exam_score_true_A17, mapping = aes(x= q1_grp)) + geom_histogram(color = "white") + labs(x = "Quiz 1 group score", y= "number of students", title = ("Quiz 1 group score Distributions by sex during A17")) + facet_wrap(~ sex_id, nrow = 3 )

#by fgn
ggplot(data = data_2s_exam_score_true_A17, mapping = aes(x= q1_grp)) + geom_histogram(color = "white") + labs(x = "Quiz 1 group score", y= "number of students", title = ("Quiz 1 group score Distributions by first generation status during A17")) + facet_wrap(~ fgn_id, nrow = 3 )

#by urm
ggplot(data = data_2s_exam_score_true_A17, mapping = aes(x= q1_grp)) + geom_histogram(color = "white") + labs(x = "Quiz 1 group score", y= "number of students", title = ("Quiz 1 group score Distributions by representation during A17")) + facet_wrap(~ urm_id, nrow = 3 )

##boxplots
ggplot(data = data_2s_exam_score_true_A17, mapping = aes(y= q1_grp)) + geom_boxplot() + labs(x = "Quiz 1 group score", y= "number of students", title = ("Quiz 1 group score Distributions by version during A17")) + facet_wrap(~ ver, nrow = 3 )

#by eop
ggplot(data = data_2s_exam_score_true_A17, mapping = aes(y= q1_grp)) + geom_boxplot() + labs(x = "Quiz 1 group score", y= "number of students", title = ("Quiz 1 group score Distributions by eop during A17")) + facet_wrap(~ eop_id, nrow = 3 )

#by sex
ggplot(data = data_2s_exam_score_true_A17, mapping = aes(y= q1_grp)) + geom_boxplot() + labs(x = "Quiz 1 group score", y= "number of students", title = ("Quiz 1 group score Distributions by sex during A17")) + facet_wrap(~ sex_id, nrow = 3 )

#by fgn
ggplot(data = data_2s_exam_score_true_A17, mapping = aes(y= q1_grp)) + geom_boxplot() + labs(x = "Quiz 1 group score", y= "number of students", title = ("Quiz 1 group score Distributions by first generation status during A17")) + facet_wrap(~ fgn_id, nrow = 3 )

#by urm
ggplot(data = data_2s_exam_score_true_A17, mapping = aes(y= q1_grp)) + geom_boxplot() + labs(x = "Quiz 1 group score", y= "number of students", title = ("Quiz 1 group score Distributions by representation during A17")) + facet_wrap(~ urm_id, nrow = 3 )
```
```{r}
ggplot(data = data_2s_exam_score_true_A17, mapping = aes(x= q1_total)) + geom_histogram(color = "white") + labs(x = "Quiz 1 total score", y= "number of students", title = ("Quiz 1 total score Distributions by version during A17")) + facet_wrap(~ ver, nrow = 3 )

#by eop
ggplot(data = data_2s_exam_score_true_A17, mapping = aes(x= q1_total)) + geom_histogram(color = "white") + labs(x = "Quiz 1 total score", y= "number of students", title = ("Quiz 1 total score Distributions by eop during A17")) + facet_wrap(~ eop_id, nrow = 3 )

#by sex
ggplot(data = data_2s_exam_score_true_A17, mapping = aes(x= q1_total)) + geom_histogram(color = "white") + labs(x = "Quiz 1 total score", y= "number of students", title = ("Quiz 1 total score Distributions by sex during A17")) + facet_wrap(~ sex_id, nrow = 3 )

#by fgn
ggplot(data = data_2s_exam_score_true_A17, mapping = aes(x= q1_total)) + geom_histogram(color = "white") + labs(x = "Quiz 1 total score", y= "number of students", title = ("Quiz 1 total score Distributions by first generation status during A17")) + facet_wrap(~ fgn_id, nrow = 3 )

#by urm
ggplot(data = data_2s_exam_score_true_A17, mapping = aes(x= q1_total)) + geom_histogram(color = "white") + labs(x = "Quiz 1 total score", y= "number of students", title = ("Quiz 1 total score Distributions by representation during A17")) + facet_wrap(~ urm_id, nrow = 3 )

##boxplots
ggplot(data = data_2s_exam_score_true_A17, mapping = aes(y= q1_total)) + geom_boxplot() + labs(x = "Quiz 1 total score", y= "number of students", title = ("Quiz 1 total score Distributions by version during A17")) + facet_wrap(~ ver, nrow = 3 )

#by eop
ggplot(data = data_2s_exam_score_true_A17, mapping = aes(y= q1_total)) + geom_boxplot() + labs(x = "Quiz 1 total score", y= "number of students", title = ("Quiz 1 total score Distributions by eop during A17")) + facet_wrap(~ eop_id, nrow = 3 )

#by sex
ggplot(data = data_2s_exam_score_true_A17, mapping = aes(y= q1_total)) + geom_boxplot() + labs(x = "Quiz 1 total score", y= "number of students", title = ("Quiz 1 total score Distributions by sex during A17")) + facet_wrap(~ sex_id, nrow = 3 )

#by fgn
ggplot(data = data_2s_exam_score_true_A17, mapping = aes(y= q1_total)) + geom_boxplot() + labs(x = "Quiz 1 total score", y= "number of students", title = ("Quiz 1 total score Distributions by first generation status during A17")) + facet_wrap(~ fgn_id, nrow = 3 )

#by urm
ggplot(data = data_2s_exam_score_true_A17, mapping = aes(y= q1_total)) + geom_boxplot() + labs(x = "Quiz 1 total score", y= "number of students", title = ("Quiz 1 total score Distributions by representation during A17")) + facet_wrap(~ urm_id, nrow = 3 )
```
#Now I am done with the graphs, I will come back and make the box plots look prettier


