---
title: "Factor Analysis exploration"
author: "ChemistryR"
date: "2023-04-19"
output: html_document
---

Loading Library
```{r}
library(tidyverse)
library(here)
library(skimr)
library(moderndive)
library(lme4)
library(MASS)
library(nnet)
library(Hmisc)
library(knitr)
library(janitor)
library(VGAM)
library(MASS)
library(ggplot2)
library(reshape2)
```

## Set up

## Designate working directories and copy file

```{r}
proj_dir <- here()
original_data_dir   <- here("original-data")
importable_data_dir <- here("processing-and-analysis", "01-importable-data")
analysis_data_dir   <- here("processing-and-analysis", "03-analysis-data")
metadata_dir <- here("original-data", "metadata")
copy_from <- paste0(original_data_dir, "two_stage_master_wide_deid.rds")
copy_to <- paste0(importable_data_dir, "two_stage_master_wide_deid.rds")
file.copy(copy_from, copy_to)
```

## Import dataset

```{r}
master_original_1 <- readRDS(here(importable_data_dir, "two_stage_master_wide_deid.rds"))
```

## Clean the dataset to be what we need

```{r}
master_two_stage <- master_original_1 %>% 
  dplyr::select(
  "two_stage_id",
  "sis_id",
  "class.x",
  "class.y",
  "exp",
  "sex_id",
  "urm_id",
  "eop_id",
  "fgn_id",
  "satm",
  "satv",
  "hs_gpa",
  "aleksikc_score",
  contains("perts_q")
  )
```

## Clean to original and final data
```{r}
survey_result_original <- master_two_stage %>% 
  dplyr::select(
  contains("1"),
  contains("exp")
  )

survey_result_final <- master_two_stage %>% 
  dplyr::select(
  contains("2"),
  contains("exp")
  )

survey_result_original <- survey_result_original[,!names(survey_result_original) %in%
                                               c("perts_q1effa2", "perts_q1relevplan2",
                                                 "perts_q1belong2", "perts_q1msfixed2")]

survey_result_final <- survey_result_final[,!names(survey_result_final) %in%
                                               c("perts_q2efftest1", "perts_q2relevskills1",
                                              "perts_q2belongrespect1","perts_q2msfixed1")]
```

Rename the data original
```{r}
names(survey_result_original) = gsub(pattern = "perts_", replacement = "", x =
                                       names(survey_result_original))
names(survey_result_original) = gsub(pattern = "msfixed", replacement = "msf", x =
                                       names(survey_result_original))
names(survey_result_original) = gsub(pattern = "belong", replacement = "bl", x =
                                       names(survey_result_original))
names(survey_result_original) = gsub(pattern = "relev", replacement = "rl", x =
                                       names(survey_result_original))
names(survey_result_original) = gsub(pattern = "respect", 
                                  replacement = "", x = names(survey_result_original))
names(survey_result_original) = gsub(pattern = "comfort", 
                                  replacement = "", x = names(survey_result_original))
names(survey_result_original) = gsub(pattern = "work", 
                                  replacement = "", x = names(survey_result_original))
names(survey_result_original) = gsub(pattern = "myself", 
                                  replacement = "", x = names(survey_result_original))
names(survey_result_original) = gsub(pattern = "plan", 
                                  replacement = "", x = names(survey_result_original))
names(survey_result_original) = gsub(pattern = "skills", 
                                  replacement = "", x = names(survey_result_original))
names(survey_result_original) = gsub(pattern = "future", 
                                  replacement = "", x = names(survey_result_original))
```

Rename the data final
```{r}
names(survey_result_final) = gsub(pattern = "perts_", replacement = "", x = 
                                    names(survey_result_final))
names(survey_result_final) = gsub(pattern = "msfixed", replacement = "msf", x =
                                       names(survey_result_final))
names(survey_result_final) = gsub(pattern = "belong", replacement = "bl", x =
                                       names(survey_result_final))
names(survey_result_final) = gsub(pattern = "relev", replacement = "rl", x =
                                       names(survey_result_final))
names(survey_result_final) = gsub(pattern = "respect", 
                                  replacement = "", x = names(survey_result_final))
names(survey_result_final) = gsub(pattern = "comfort", 
                                  replacement = "", x = names(survey_result_final))
names(survey_result_final) = gsub(pattern = "work", 
                                  replacement = "", x = names(survey_result_final))
names(survey_result_final) = gsub(pattern = "myself", 
                                  replacement = "", x = names(survey_result_final))
names(survey_result_final) = gsub(pattern = "plan", 
                                  replacement = "", x = names(survey_result_final))
names(survey_result_final) = gsub(pattern = "skills", 
                                  replacement = "", x = names(survey_result_final))
names(survey_result_final) = gsub(pattern = "future", 
                                  replacement = "", x = names(survey_result_final))
```

original-control and experiment
```{r}
survey_result_original_control <-filter(survey_result_original, exp == "CONTROL")
survey_result_original_experiment <-filter(survey_result_original, exp == "EXPERIMENTAL")
survey_result_original_control <- survey_result_original_control [,!names(survey_result_original_control)
                                                                  %in% c("exp")]
survey_result_original_experiment <- survey_result_original_experiment [,!names(survey_result_original_experiment)
                                                                  %in% c("exp")]
colnames(survey_result_original_experiment) <- paste(colnames(survey_result_original_experiment),"E",sep="")
colnames(survey_result_original_control) <- paste(colnames(survey_result_original_control),"C",sep="")
```

final control and experiment
```{r}
survey_result_final_control <-filter(survey_result_final, exp == "CONTROL")
survey_result_final_experiment <-filter(survey_result_final, exp == "EXPERIMENTAL")
survey_result_final_control <- survey_result_final_control [,!names(survey_result_final_control)
                                                                  %in% c("exp")]
survey_result_final_experiment <- survey_result_final_experiment [,!names(survey_result_final_experiment)
                                                                  %in% c("exp")]
colnames(survey_result_final_experiment) <- paste(colnames(survey_result_final_experiment),"E",sep="")
colnames(survey_result_final_control) <- paste(colnames(survey_result_final_control),"C",sep="")
```

## Factor the Survey Data
```{r}
lvl <- c("Strongly Disagree", "Disagree", 
        "Somewhat Disagree","Somewhat Agree", 
        "Agree", "Strongly Agree")
survey_result_orginal_experiment <- lapply(survey_result_original_experiment, factor, levels = lvl)
survey_result_final_experiment <- lapply(survey_result_final_experiment, factor, levels = lvl)
survey_result_orginal_control <- lapply(survey_result_original_control, factor, levels = lvl)
survey_result_final_control <- lapply(survey_result_final_control, factor, levels = lvl)
```
order factor the data
```{r}

survey_result_orginal_experiment <- factor(c(lvl), 
                                           levels = lvl, 
                                           ordered = TRUE)
survey_result_final_experiment <- factor(c(lvl), 
                                           levels = lvl, 
                                           ordered = TRUE)
survey_result_final_control <- factor(c(lvl), 
                                           levels = lvl, 
                                           ordered = TRUE)
survey_result_orginal_control <- factor(c(lvl), 
                                           levels = lvl, 
                                           ordered = TRUE)
## Data Summary

glimpse(survey_result_orginal_control)
glimpse(survey_result_final_experiment)
glimpse(survey_result_orginal_control)
glimpse(survey_result_final_control)
```

## First Try

survey_result_year_1_plot_n <- data.frame(survey_result_year_1)
survey_result_year_1_plot_n$question <- row.names(survey_result_year_1_plot_n)
library(reshape2)
melt_survey_result_year_1_plot_n <- melt(survey_result_year_1_plot_n,"question")
library(ggplot2)
survey_result_year_1_plot_n <- ggplot(melt_survey_result_year_1_plot_n,
                                    aes(x = variable, fill = value))
survey_result_year_1_plot_n + geom_dotplot(stackgroups = TRUE,binwidth = 0.1)



## Change Factor variable to numerical variable
```{r}
survey_result_orginal_experiment <- lapply(survey_result_orginal_experiment, as.numeric)
survey_result_year_2 <- lapply(survey_result_year_2, as.numeric)
```


## Data Summary

glimpse(survey_result_year_1)
glimpse(survey_result_year_2)

## Second Try

survey_result_year_1_plot = data.frame(survey_result_year_1)
plot(survey_result_year_1_plot, aes(variable,value))


## Third Try

column_survey_result_year_1<-melt(survey_result_year_1)
ggplot(column_survey_result_year_1,aes(L1,value))+geom_point()


## Forth Try

library(plot.matrix)
matplot(as.data.frame(survey_result_year_1), type = "l")


## fifth Try

library(psych)
survey_result_year_1 <-data.frame(survey_result_year_1)
head(survey_result_year_1)
plot(survey_result_year_1, pch = 1, cex = 0.2)


## Six Try: Individually

msfix1_survey_result <- survey_result_year_1 %>% 
  dplyr::select(
    contains("msfixed")
  )
library(psych)
msfix1_survey_result <-data.frame(msfix1_survey_result)
head(msfix1_survey_result)
plot(msfix1_survey_result, pch = 10, cex = 2)



belong_survey_result <- survey_result_year_1 %>% 
  dplyr::select(
    contains("belong")
  )
library(psych)
msfix1_survey_result <-data.frame(belong_survey_result)
head(belong_survey_result)
plot(belong_survey_result, pch = 10, cex = 2)



rele_survey_result <- survey_result_year_1 %>% 
  dplyr::select(
    contains("rele")
  )
library(psych)
msfix1_survey_result <-data.frame(rele_survey_result)
head(rele_survey_result)
plot(rele_survey_result, pch = 10, cex = 2)


eff_survey_result <- survey_result_year_1 %>% 
  dplyr::select(
    contains("eff")
  )
library(psych)
eff_survey_result <-data.frame(eff_survey_result)
head(eff_survey_result)
plot(eff_survey_result, pch = 10, cex = 2)


## create_data_table for experiment final
```{r}
column_survey_result_final_experiment <- melt(survey_result_final_experiment)
colnames(column_survey_result_final_experiment)[2] = "question_name"
column_survey_result_final_experiment %>% group_by(question_name) %>% tally() %>% arrange(desc(n))
pivot_long_column_survey_result_final_experiment <- column_survey_result_final_experiment %>%
  group_by(question_name, value) %>% tally() %>% 
  arrange(desc(n))
pivot_long_column_survey_result_final_experiment
```
## create_data_table for control final
```{r}
column_survey_result_final_control <- melt(survey_result_final_control)
colnames(column_survey_result_final_control)[2] = "question_name"
column_survey_result_final_control %>% group_by(question_name) %>% tally() %>% arrange(desc(n))
pivot_long_column_survey_result_final_control <- column_survey_result_final_control %>%
  group_by(question_name, value) %>% tally() %>% 
  arrange(desc(n))
pivot_long_column_survey_result_final_control
```
## create_data_table for control original
```{r}
column_survey_result_original_control <- melt(survey_result_orginal_control)
colnames(column_survey_result_original_control)[2] = "question_name"
column_survey_result_original_control %>% group_by(question_name) %>% tally() %>% arrange(desc(n))
pivot_long_column_survey_result_original_control <- column_survey_result_original_control %>%
  group_by(question_name, value) %>% tally() %>% 
  arrange(desc(n))
pivot_long_column_survey_result_original_control
```

## create_data_table for experiment original
```{r}
column_survey_result_original_experiment <- melt(survey_result_orginal_experiment)
colnames(column_survey_result_original_experiment)[2] = "question_name"
column_survey_result_original_experiment %>% group_by(question_name) %>% tally() %>% arrange(desc(n))
pivot_long_column_survey_result_original_experiment <- column_survey_result_original_experiment %>%
  group_by(question_name, value) %>% tally() %>% 
  arrange(desc(n))
pivot_long_column_survey_result_original_experiment
```

Graph experiment final
```{r}
ggplot(data = subset(pivot_long_column_survey_result_final_experiment, !is.na(value)),
       aes(question_name,value)) + 
  geom_point(aes(size = n)) +
  scale_size_continuous(range = c(1, 8))
```

Graph control final
```{r}
ggplot(data = subset(pivot_long_column_survey_result_final_control, !is.na(value)),
       aes(question_name,value)) + 
  geom_point(aes(size = n)) +
  scale_size_continuous(range = c(1, 8))
```

Graph control original
```{r}
ggplot(data = subset(pivot_long_column_survey_result_original_control, !is.na(value)),
       aes(question_name,value)) + 
  geom_point(aes(size = n)) +
  scale_size_continuous(range = c(1, 8))
```

Graph experiment original
```{r}
ggplot(data = subset(pivot_long_column_survey_result_original_experiment, !is.na(value)),
       aes(question_name,value)) + 
  geom_point(aes(size = n)) +
  scale_size_continuous(range = c(1, 8))
```

data comparison:
```{r}
year_con <- rbind(pivot_long_column_survey_result_original_control,
                  pivot_long_column_survey_result_final_control)
year_con <- arrange(year_con, desc(question_name))
year_exp <- rbind(pivot_long_column_survey_result_original_experiment,
                  pivot_long_column_survey_result_final_experiment)
year_exp <- arrange(year_exp, desc(question_name))
```

Control year
```{r}
ggplot(data = subset(year_con, !is.na(value)),
       aes(question_name,value)) + 
  geom_point(aes(size = n)) +
  scale_size_continuous(range = c(0.5, 3))
```

Experiment year
```{r}
ggplot(data = subset(year_exp, !is.na(value)),
       aes(question_name,value)) + 
  geom_point(aes(size = n)) +
  scale_size_continuous(range = c(0.5, 3))
```


Total comparison
```{r}
con_exp <- rbind(year_con, year_exp)
con_exp <- arrange(con_exp, desc(question_name))
```
Comparison graph
```{r}
ggplot(data = subset(con_exp, !is.na(value)),
       aes(question_name,value)) + 
  geom_point(aes(size = n)) +
  scale_size_continuous(range = c(0.2, 2))
```
