---
title: "Model_Analysis_New_Data"
author: "Ganling"
date: "8/30/2021"
output: 
  html_document:
    keep_md: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(fs) # file system operations functions
library(readxl)
library(janitor)
library(knitr)
library(lme4)
```
## Set working directories
```{r}
original_data_dir   <- here("original-data", "/")
importable_data_dir <- here("processing-and-analysis", "01-importable-data", "/")
analysis_data_dir   <- here("processing-and-analysis", "03-analysis-data", "/")
```

## Import data set
### Copy data from `original-data` to `01-importable-data`
```{r}
copy_from <- paste0(original_data_dir, 
 "two_stage_master_wide_deid.rds")
copy_to <- paste0(importable_data_dir, "two_stage_master_wide_deid.rds")
file.copy(copy_from, copy_to)
```
### Import master data set
```{r}
df <- readRDS(paste0(importable_data_dir, "two_stage_master_wide_deid.rds"))
```
## Model Selection
### Change `NAs` to experiment and select columns
```{r}
df_1 <- df %>%
  mutate(exp = replace_na(exp, "EXPERIMENT"))
df_true <- df_1 %>%
  select(two_stage_id, class.x, class.y, qtr, course_fullid, ta_sect, exp, exam1_c, exam2_c, final_c, satm_c, satv_c, aleksikc_c, hs_gpa_c, final_c, sex_id, urm_id, eop_id, fgn_id)
```
### Model selection on `sex_id`
#### Sex Model 1: fixed effect
```{r}
sex_mod1 <- lm(final_c ~ exp * sex_id + 
                 exp + sex_id + 
                 satm_c + satv_c + aleksikc_c + hs_gpa_c, 
               data = df_true)
summary(sex_mod1)
```
#### Sex model 2: add in random effect
```{r}
sex_mod2 <- lmer(final_c ~ exp * sex_id + 
                   exp + sex_id +
                   satm_c + satv_c + aleksikc_c + hs_gpa_c + 
                   (1 | ta_sect),
                 data = df_true, REML = TRUE)
summary(sex_mod2)

AIC(sex_mod1, sex_mod2)
```
* Since the AIC value went up after adding random effect, we can ignore the random effect. 

#### Sex model 3: no interaction or random effect
```{r}
sex_mod3 <- lm(final_c ~ exp + sex_id + 
                 satm_c + satv_c + aleksikc_c + hs_gpa_c, 
               data = df_true)
AIC(sex_mod1, sex_mod3)
summary(sex_mod3)
```
* The AIC value decreased for `sex_mod3`
* The interaction between `exp` and `sex_id` should not be included

#### Sex model 3.a: No `exp` 
```{r}
sex_mod3.a <- lm(final_c ~ sex_id + 
                 satm_c + satv_c + aleksikc_c + hs_gpa_c, 
               data = df_true)
AIC(sex_mod3, sex_mod3.a)
summary(sex_mod3.a)
```
* The AIC value slightly decreased
* Since sex_mod3.a is slightly smaller, we keep this model

#### Sex model 3.b: Remove `sex_id`
```{r}
sex_mod3.b <- lm(final_c ~ satm_c + satv_c + aleksikc_c + hs_gpa_c, 
               data = df_true)
AIC(sex_mod3.a, sex_mod3.b)
summary(sex_mod3.b)
```
* Since the AIC value increased, `sex_id` should be kept in the model

#### Sex model 3.c: Remove `satv_c`
```{r}
sex_mod3.c <- lm(final_c ~ sex_id + 
                 satm_c + aleksikc_c + hs_gpa_c, 
               data = df_true)
AIC(sex_mod3.a, sex_mod3.c)
summary(sex_mod3.c)
```
* The AIC value went up, so `satv_c` should be kept in our model 

#### Sex model 3.d: Remove `aleksikc_c`
```{r}
sex_mod3.d <- lm(final_c ~ sex_id + 
                 satm_c + satv_c + hs_gpa_c, 
               data = df_true)
AIC(sex_mod3.a, sex_mod3.d)
summary(sex_mod3.d)
```
* The AIC value increased, `aleksikc_c` should be kept

#### Sex Model 3.e: remove `hs_gpa_c` 
```{r}
sex_mod3.e <- lm(final_c ~ sex_id + 
                 satm_c + satv_c + aleksikc_c , 
               data = df_true)
AIC(sex_mod3.a, sex_mod3.e)
summary(sex_mod3.e)
```
* The AIC value went up, `hs_gpa_c` is kept

#### Sex Model 3.f: Remove `satm_c`
```{r}
sex_mod3.f <- lm(final_c ~ sex_id + 
                 satv_c + aleksikc_c + hs_gpa_c, 
               data = df_true)
AIC(sex_mod3.a, sex_mod3.f)
summary(sex_mod3.f)
```
* Again, the AIC value went up. `satm_c` should be kept
* `sex_mod3.a` is the best fitting model  

### Model selection on `eop_id`
#### Eop model 1: fixed effect
```{r}
eop_mod1 <- lm(final_c ~ exp * eop_id + 
                 exp + eop_id + 
                 satm_c + satv_c + aleksikc_c + hs_gpa_c, 
               data = df_true)
summary(eop_mod1)
```
#### EOP model 2: random effect
```{r}
eop_mod2 <- lmer(final_c ~ exp * eop_id + 
                   exp + eop_id +
                   satm_c + satv_c + aleksikc_c + hs_gpa_c + 
                   (1 | ta_sect),
                 data = df_true, REML = TRUE)
summary(eop_mod2)
AIC(eop_mod1, eop_mod2)
```
* AIC value went up, random effect should not be removed
* Interaction has the lowest t value, it should be removed next

#### Eop model 3: remove `exp` and `eop_id` interaction and no radnom effect 
```{r}
eop_mod3 <- lm(final_c ~ exp + eop_id + 
                 satm_c + satv_c + aleksikc_c + hs_gpa_c, 
               data = df_true)
AIC(eop_mod1, eop_mod3)
summary(eop_mod3)
```
* The AIC value decreased with the new model, meaning the interaction should be removed


#### eop model 3.a: remove `exp`
```{r}
eop_mod3.a <- lm(final_c ~ eop_id + satm_c + satv_c + aleksikc_c + hs_gpa_c, 
               data = df_true)
AIC(eop_mod3, eop_mod3.a)
summary(eop_mod3.a)
```
* The AIC value decreased, `exp` should be removed from the model 

#### eop model 3.b: remove `eop_id`
```{r}
eop_mod3.b <- lm(final_c ~ satm_c + satv_c + aleksikc_c + hs_gpa_c, 
               data = df_true)
AIC(eop_mod3.a, eop_mod3.b)
summary(eop_mod3.b)
```
* The AIC value decreased slightly, and the `eop_mod3.b` has the lower value. `eop_id` should be removed

#### eop model 3.c: remove `satv_c`
```{r}
eop_mod3.c <- lm(final_c ~ satm_c + aleksikc_c + hs_gpa_c, 
               data = df_true)
AIC(eop_mod3.b, eop_mod3.c)
summary(eop_mod3.c)
```
* The AIC value went up, `satv_c` should be kept. 
* `eop_mod3.b` is the best fitting model

### Model selection on `fgn_id`
#### Fgn model 1: fixed effect
```{r}
fgn_mod1 <- lm(final_c ~ exp * fgn_id + 
                 exp + fgn_id + 
                 satm_c + satv_c + aleksikc_c + hs_gpa_c, 
               data = df_true)
summary(fgn_mod1)
```
#### Fgn model 2: Random effect
```{r}
fgn_mod2 <- lmer(final_c ~ exp * fgn_id + 
                   exp + fgn_id +
                   satm_c + satv_c + aleksikc_c + hs_gpa_c + 
                   (1 | ta_sect),
                 data = df_true, REML = TRUE)
summary(fgn_mod2)
AIC(fgn_mod1,fgn_mod2)
```
* AIC value went up, random effect should not be included

#### Fgn model 3: remove `exp` and `fgn_id` interaction and no radnom effect 
```{r}
fgn_mod3 <- lm(final_c ~ exp + fgn_id + 
                 satm_c + satv_c + aleksikc_c + hs_gpa_c, 
               data = df_true)
AIC(fgn_mod1, fgn_mod3)
summary(fgn_mod3)
```
* The AIC value went up, the demographic `interaction` should be retained 

#### Fgn model 3.a: remove `fgn_id`
```{r}
fgn_mod3.a <- lm(final_c ~ exp + 
                 satm_c + satv_c + aleksikc_c + hs_gpa_c, 
               data = df_true)
AIC(fgn_mod1, fgn_mod3.a)
summary(fgn_mod3.a)
```
*The AIC value went up, suggesting fgn_mod1 is the best fitting model 
### Model selection on `urm_id`
#### Urm model 1: fixed effect 
```{r}
urm_mod1 <- lm(final_c ~ exp * urm_id + 
                 exp + urm_id + 
                 satm_c + satv_c + aleksikc_c + hs_gpa_c, 
               data = df_true)
summary(urm_mod1)
```
#### Urm model 2: random effect
```{r}
urm_mod2 <- lmer(final_c ~ exp * urm_id + 
                   exp + urm_id +
                   satm_c + satv_c + aleksikc_c + hs_gpa_c + 
                   (1 | ta_sect),
                 data = df_true, REML = TRUE)
summary(urm_mod2)
AIC(urm_mod1, urm_mod2)
```
* The AIC value went up, random effect should not be kept
* Next, `interaction` should be removed

#### Urm model 3: Remove random effects and interactions
```{r}
urm_mod3 <- lm(final_c ~ exp + urm_id + 
                 satm_c + satv_c + aleksikc_c + hs_gpa_c, 
               data = df_true)
AIC(urm_mod1, urm_mod3)
summary(urm_mod3) 
```
* The AIC value went down, thus the interaction should be removed

#### urm model 3.a: Remove `urm_id`
```{r}
urm_mod3.a <- lm(final_c ~ exp + 
                 satm_c + satv_c + aleksikc_c + hs_gpa_c, 
               data = df_true)
AIC(urm_mod3, urm_mod3.a)
summary(urm_mod3.a)
```
*The AIC value went up, `urm_id` should be retained

#### urm model 3.b: remove `exp`
```{r}
urm_mod3.b <- lm(final_c ~ urm_id +
                   satm_c + satv_c + aleksikc_c + hs_gpa_c, 
               data = df_true)
AIC(urm_mod3, urm_mod3.b)
summary(urm_mod3.b)
```
* The AIC value decreased, so `exp` should be removed 
** The T value is larger for `exp` compared to `urm_id` but the AIC value decreased only when we removed `exp`

#### Urm model 3.c: Remove `satv_c`
```{r}
urm_mod3.c <- lm(final_c ~ urm_id +
                   satm_c + aleksikc_c + hs_gpa_c, 
               data = df_true)
AIC(urm_mod3.b, urm_mod3.c)
```
* The AIC value went up, `urm_mod3.b` is the best fitting model
















