---
title: "TA Section Random Effects"
author: "Ganling"
date: "5/9/2021"
output: 
  html_document:
    keep_md: yes
---
```{r}
library(here)
library(tidyverse)
library(skimr)
library(moderndive)
library(ggplot2)
library(dplyr)
library(readr)
library(lme4)
```
### Set working directories
```{r}
proj_dir <- here()
#add original original data directory sting
original_data_dir <- here("original-data", "/")
#add numbers to importable and analysis data directory strings
importable_data_dir <- here("processing-and-analysis", "01-importable-data", "/")
#add analysis directory string
analysis_data_dir <- here("processing-and-analysis", "03-analysis-data", "/")
metadata_dir <- here("original-data", "metadata", "/")
```
### load the data set through here directory
```{r}
# CFC: Copy rds file from the `original_data_dir` to the `importable_data_dir`.
copy_from <- paste0(original_data_dir, "master_2s_small_deid_scaled")
copy_to <- paste0(importable_data_dir, "master_2s_small_deid_scaled")
# CFC: If the file already exists in the target directory, the file will not be 
# overwritten and this command will return `FALSE`.
file.copy(copy_from, copy_to)
```
### Import and read the data
```{r}
#data_2s <- readRDS(paste0(metadata_dir, "master_2s_small_deid_scaled.rds"))
data_2s <- readRDS("G:/Shared drives/CEG Two-Stage Exams Analysis/Ganling/two-stage-2/original-data/master_2s_small_deid_scaled.rds")
```
## Data wrangling
#### Creating unique rows and remove the ones that are not necessary for this data model
```{r}
data_2s_extras_removed = subset(data_2s, select = -c(item_num, stud_ans, exam_key, corr, bloom_rating, complexity_rating_mean, complexity_rating_median, item_code, stem))
data_2s_unique = unique(data_2s_extras_removed, incomparables = FALSE)
View(data_2s_unique)
data_2s_exam_score <- data_2s_unique %>% 
  select(exam1, exam2, finalexam, course_fullid, ver, sex_id, urm_id, eop_id, fgn_id, experiment1, ta_sect, satverbal, satmath, high_sch_gpa, mastered_topics_initial_kc)
  
data_2s_exam_score_true <- na.omit(data_2s_exam_score)
```
## Make 'ta_sect'unique across different data values
#### First, we need to split the data from into year 16 and year 17 
```{r}
data_2s_exam_score_true$ta_sect <- as.character(data_2s_exam_score_true$ta_sect)

#next, split the dataframe in half
master_unique_2016 <- subset(data_2s_exam_score_true, course_fullid != "CHEM_142_A_2017_4" & course_fullid != "CHEM_142_B_2017_4")
master_unique_2017 <- subset(data_2s_exam_score_true, course_fullid != "CHEM_142_A_2016_4" & course_fullid != "CHEM_142_B_2016_4")
```
#### Use the vectorization function in R, we can add `_16` and `_17`at the end of each TA section 
```{r}
master_unique_2016$ta_sect <- paste0(master_unique_2016$ta_sect, "_16")
master_unique_2017$ta_sect <- paste0(master_unique_2017$ta_sect, "_17")
```
## recombine the dataframes together and make the `ta_sect` factors again
```{r}
master_true <- rbind(master_unique_2016, master_unique_2017)
master_true$ta_sect <- as.factor(master_true$ta_sect)
```
#### I would like to create a model that could implement the random effects based on the TA sections
```{r}
#First, include sex_id as a control variable
mod1 <- lmer(finalexam ~ experiment1 + satmath + satverbal + mastered_topics_initial_kc + high_sch_gpa + sex_id + (1|ta_sect), 
           data=master_true)
summary(mod1)
AIC(mod1)

#Include eop_id as a control variable
mod2 <- lmer(finalexam ~ experiment1 + satmath + satverbal + mastered_topics_initial_kc + high_sch_gpa + eop_id + (1|ta_sect), 
           data=master_true)
summary(mod2)
AIC(mod2)

#Include fgn_id as a control variable
mod3 <- lmer(finalexam ~ experiment1 + satmath + satverbal + mastered_topics_initial_kc + high_sch_gpa + fgn_id + (1|ta_sect), 
           data=master_true)
summary(mod3)
AIC(mod3)

#Include urm_id as a control variable
mod4 <- lmer(finalexam ~ experiment1 + satmath + satverbal + mastered_topics_initial_kc + high_sch_gpa + urm_id + (1|ta_sect), 
           data=master_true)
summary(mod4)
AIC(mod4)
#
```
#### This is what I got previously by changing out different independent variables, now I wnat to compare only the models with and without random effects
```{r}
#First, test that the treatment impacts scores
mod1 <- lm(finalexam ~ experiment1 + satmath + satverbal + mastered_topics_initial_kc + high_sch_gpa, 
           data=master_true)
summary(mod1)
AIC(mod1)

#Include random effects (clustering)
mod2 <- lmer(finalexam ~ experiment1 + satmath + satverbal + mastered_topics_initial_kc + high_sch_gpa + (1|ta_sect), 
           data=master_true)
summary(mod2)
AIC(mod2)
```
```{r}
AIC(mod1, mod2)
```
#### Again, the change of AIC is much bigger than 2




