---
title: "TA_Random_Effects_Interaction"
author: "Ganling"
date: "5/9/2021"
output: 
  html_document:
    keep_md: yes
---
### Setting up the data set 
#### Read libraries 
```{r}
library(here)
library(tidyverse)
library(fs) # file system operations functions
library(readxl)
library(janitor)
library(knitr)
library(lme4)
```
#### Load the data set
```{r}
original_data_dir   <- here("original-data", "/")
importable_data_dir <- here("processing-and-analysis", "01-importable-data", "/")
analysis_data_dir   <- here("processing-and-analysis", "03-analysis-data", "/")

df <- readRDS("G:/Shared drives/CEG Two-Stage Exams Analysis/Ganling/two-stage-2/original-data/master_2s_small_deid_scaled.rds")
```
### data wrangling 
#### creating data frames with unique `ta_sect` 
```{r}
df2 <- df %>% select(course_fullid)

df3 <- df2 %>% mutate(year = str_extract(course_fullid, "\\d\\d\\d\\d")) 

df4 <- df3 %>% 
  mutate(year = ifelse(str_detect(course_fullid, "2016"), 
                       paste(year, "16", sep = "_"),
                       paste(year, "17", sep = "_")))
```
#### NOTE: I am trying to add A and B after the year, but I can't seem to make it work 
#### Now using my orginal method.Creating unique rows and remove the ones that are not necessary for this data model
```{r}
data_2s_extras_removed = subset(df, select = -c(item_num, stud_ans, exam_key, corr, bloom_rating, complexity_rating_mean, complexity_rating_median, item_code, stem))
data_2s_unique = unique(data_2s_extras_removed, incomparables = FALSE)
View(data_2s_unique)
data_2s_exam_score <- data_2s_unique %>% 
  select(exam1, exam2, finalexam, course_fullid, ver, sex_id, urm_id, eop_id, fgn_id, experiment1, ta_sect, satverbal, satmath, high_sch_gpa, mastered_topics_initial_kc)
  
data_2s_exam_score_true <- na.omit(data_2s_exam_score)
```
### Make `ta_sect` unique across different data values
#### First, we need to split the data from into year 16 and year 17 
```{r}
data_2s_exam_score_true$ta_sect <- as.character(data_2s_exam_score_true$ta_sect)

#next, split the dataframe in half
master_unique_2016 <- subset(data_2s_exam_score_true, course_fullid != "CHEM_142_A_2017_4" & course_fullid != "CHEM_142_B_2017_4")
master_unique_2017 <- subset(data_2s_exam_score_true, course_fullid != "CHEM_142_A_2016_4" & course_fullid != "CHEM_142_B_2016_4")
```
#### Use the vectorization function in R, we can add `_16` and `_17` at the end of each TA section 
```{r}
master_unique_2016$ta_sect <- paste0(master_unique_2016$ta_sect, "_16")
master_unique_2017$ta_sect <- paste0(master_unique_2017$ta_sect, "_17")
```
#### recombine the dataframes together and make the `ta_sect` factors again
```{r}
master_true <- rbind(master_unique_2016, master_unique_2017)
master_true$ta_sect <- as.factor(master_true$ta_sect)
```
### Interaction Models
#### Creating interaction models based on the demographics, `sex_id`, `eop_id`, `fgn_id` and `urm_id`
```{r}
# Without random effect but set `sex_id` as an interacting variable
mod1 <- lm(finalexam ~ experiment1*sex_id + satmath + satverbal + mastered_topics_initial_kc + high_sch_gpa, 
           data=master_true)
summary(mod1)
AIC(mod1)
#include `sex_id` as an interacting variable
mod2 <- lmer(finalexam ~ experiment1*sex_id + satmath + satverbal + mastered_topics_initial_kc + high_sch_gpa + (1|ta_sect),
           data=master_true)
summary(mod2)
AIC(mod2)

AIC(mod1, mod2)
```
```{r}
# Without random effect but set `eop_id` as an interacting variable
mod3 <- lm(finalexam ~ experiment1*eop_id + satmath + satverbal + mastered_topics_initial_kc + high_sch_gpa, 
           data=master_true)
summary(mod3)
AIC(mod3)
#include `sex_id` as an interacting variable
mod4 <- lmer(finalexam ~ experiment1*eop_id + satmath + satverbal + mastered_topics_initial_kc + high_sch_gpa + (1|ta_sect),
           data=master_true)
summary(mod4)
AIC(mod4)

AIC(mod3, mod4)
```
```{r}
# Without random effect but set `fgn_id` as an interacting variable
mod5 <- lm(finalexam ~ experiment1*fgn_id + satmath + satverbal + mastered_topics_initial_kc + high_sch_gpa, 
           data=master_true)
summary(mod5)
AIC(mod5)
#include `fgn_id` as an interacting variable
mod6 <- lmer(finalexam ~ experiment1*fgn_id + satmath + satverbal + mastered_topics_initial_kc + high_sch_gpa + (1|ta_sect),
           data=master_true)
summary(mod6)
AIC(mod6)

AIC(mod5, mod6)
```
```{r}
# Without random effect but set `urm_id` as an interacting variable
mod7 <- lm(finalexam ~ experiment1*urm_id + satmath + satverbal + mastered_topics_initial_kc + high_sch_gpa, 
           data=master_true)
summary(mod7)
AIC(mod7)
#include `urm_id` as an interacting variable
mod8 <- lmer(finalexam ~ experiment1*urm_id + satmath + satverbal + mastered_topics_initial_kc + high_sch_gpa + (1|ta_sect),
           data=master_true)
summary(mod8)
AIC(mod8)

AIC(mod7, mod8)
```
#### NOTE: The AIC differences are all greater than 2 when interacting with different demographics.  

### Models without unique `ta_sect`
```{r}
df5_true <- rename(data_2s_exam_score_true)

#First, test that the treatment impacts scores
mod9 <- lm(finalexam ~ experiment1 + satmath + satverbal + mastered_topics_initial_kc + high_sch_gpa, 
           data=df5_true)
summary(mod9)
AIC(mod9)

#Include random effects (clustering)
mod10 <- lmer(finalexam ~ experiment1 + satmath + satverbal + mastered_topics_initial_kc + high_sch_gpa + (1|ta_sect), 
           data=df5_true)
summary(mod10)
AIC(mod10)

```
```{r}
AIC(mod9,mod10)
```
#### Values with unique `ta_sect` are: 	
df
<dbl>
AIC
<dbl>
mod1 (linear fit)	7	63734.08		
mod2 (random effect)	63561.45

#### NOTE: When considering random effects with unique `ta_sect`, the AIC value decreases compared to the AIC value without unique `ta_sect`

## Model Secletion
#### Finding the best fit model for `sex_id`
```{r}
# Making a model that has fixed effects only with `lm()`
sex_mod1 <- lm(finalexam ~ experiment1 * sex_id + experiment1 + sex_id + satmath + satverbal + mastered_topics_initial_kc +  high_sch_gpa, data = master_true)

# Add in random effects for `ta_sect`
sex_mod2 <- lmer(finalexam ~ experiment1 * sex_id + experiment1 + sex_id + satmath + satverbal + mastered_topics_initial_kc +  high_sch_gpa + (1 | ta_sect), data = master_true, REML = TRUE)

AIC(sex_mod1, sex_mod2)

# remove experiment1 * dem_id from the best model predicted 
sex_mod3 <- lmer(finalexam ~ experiment1 + sex_id + satmath + satverbal + mastered_topics_initial_kc +  high_sch_gpa + (1 | ta_sect), data = master_true, REML = FALSE)

AIC(sex_mod2, sex_mod3)

#Refit model 3 with lmer and REML = FALSE
sex_mod4 <- lmer(finalexam ~ experiment1 + sex_id + satmath + satverbal + mastered_topics_initial_kc +  high_sch_gpa + (1 | ta_sect), data = master_true, REML = TRUE)

AIC(sex_mod3, sex_mod4)

```
#### Finding the best fit model for `eop_id`
```{r}
# Making a model that has fixed effects only with `lm()`
eop_mod1 <- lm(finalexam ~ experiment1 * eop_id + experiment1 + eop_id + satmath + satverbal + mastered_topics_initial_kc +  high_sch_gpa, data = master_true)

# Add in random effects for `ta_sect`
eop_mod2 <- lmer(finalexam ~ experiment1 * eop_id + experiment1 + eop_id + satmath + satverbal + mastered_topics_initial_kc +  high_sch_gpa + (1 | ta_sect), data = master_true, REML = TRUE)

AIC(eop_mod1, eop_mod2)

# remove experiment1 * dem_id from the best model predicted 
eop_mod3 <- lmer(finalexam ~ experiment1 + eop_id + satmath + satverbal + mastered_topics_initial_kc +  high_sch_gpa + (1 | ta_sect), data = master_true, REML = FALSE)

AIC(eop_mod2, eop_mod3)

#Refit model 3 with lmer and REML = FALSE
eop_mod4 <- lmer(finalexam ~ experiment1 + eop_id + satmath + satverbal + mastered_topics_initial_kc +  high_sch_gpa + (1 | ta_sect), data = master_true, REML = TRUE)

AIC(eop_mod3, eop_mod4)

```
#### Finding the best fit model for `fgn_id`
```{r}
# Making a model that has fixed effects only with `lm()`
fgn_mod1 <- lm(finalexam ~ experiment1 * fgn_id + experiment1 + fgn_id + satmath + satverbal + mastered_topics_initial_kc +  high_sch_gpa, data = master_true)

# Add in random effects for `ta_sect`
fgn_mod2 <- lmer(finalexam ~ experiment1 * fgn_id + experiment1 + fgn_id + satmath + satverbal + mastered_topics_initial_kc +  high_sch_gpa + (1 | ta_sect), data = master_true, REML = TRUE)

AIC(fgn_mod1, fgn_mod2)

# remove experiment1 * dem_id from the best model predicted 
fgn_mod3 <- lmer(finalexam ~ experiment1 + fgn_id + satmath + satverbal + mastered_topics_initial_kc +  high_sch_gpa + (1 | ta_sect), data = master_true, REML = FALSE)

AIC(fgn_mod2, fgn_mod3)

#Refit model 3 with lmer and REML = FALSE
fgn_mod4 <- lmer(finalexam ~ experiment1 + fgn_id + satmath + satverbal + mastered_topics_initial_kc +  high_sch_gpa + (1 | ta_sect), data = master_true, REML = TRUE)

AIC(fgn_mod3, fgn_mod4)

```
#### Finding the best fit model for `urm_id`
```{r}
# Making a model that has fixed effects only with `lm()`
urm_mod1 <- lm(finalexam ~ experiment1 * urm_id + experiment1 + urm_id + satmath + satverbal + mastered_topics_initial_kc +  high_sch_gpa, data = master_true)

# Add in random effects for `ta_sect`
urm_mod2 <- lmer(finalexam ~ experiment1 * urm_id + experiment1 + fgn_id + satmath + satverbal + mastered_topics_initial_kc +  high_sch_gpa + (1 | ta_sect), data = master_true, REML = TRUE)

AIC(urm_mod1, urm_mod2)

# remove experiment1 * dem_id from the best model predicted 
urm_mod3 <- lmer(finalexam ~ experiment1 + urm_id + satmath + satverbal + mastered_topics_initial_kc +  high_sch_gpa + (1 | ta_sect), data = master_true, REML = FALSE)

AIC(fgn_mod2, fgn_mod3)

#Refit model 3 with lmer and REML = FALSE
urm_mod4 <- lmer(finalexam ~ experiment1 + urm_id + satmath + satverbal + mastered_topics_initial_kc +  high_sch_gpa + (1 | ta_sect), data = master_true, REML = TRUE)

AIC(urm_mod3, urm_mod4)
```
#### For each demographic identifier, the best fitting model is the one with random effects and without interactions (model 3). The AIC value decreases by adding random effects and removing interaction. 









