---
title: "random_effect_model_selection"
author: "Ganling"
date: "8/15/2021"
output: 
  html_document:
    keep_md: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(fs) # file system operations functions
library(readxl)
library(janitor)
library(knitr)
library(lme4)
```

## Set working directories
```{r}
original_data_dir   <- here("original-data", "/")
importable_data_dir <- here("processing-and-analysis", "01-importable-data", "/")
analysis_data_dir   <- here("processing-and-analysis", "03-analysis-data", "/")
```

## Import data set
### Copy data from `original-data` to `01-importable-data`
```{r}
copy_from <- paste0(original_data_dir, "master_2s_small_deid_scaled.rds")
copy_to <- paste0(importable_data_dir, "master_2s_small_deid_scaled.rds")
file.copy(copy_from, copy_to)
```
### Import master data set
```{r}
df <- readRDS(paste0(importable_data_dir, "master_2s_small_deid_scaled.rds"))
```

## Data wrangling 
### Creating unique rows and omit NA's
```{r}
df_removed <- subset(df, select = -c(item_num, stud_ans, exam_key, corr, bloom_rating, complexity_rating_mean, complexity_rating_median, item_code, stem))
df_unique_1 <- df_removed %>% 
select(two_stage_id, exam1, exam2, finalexam, course_fullid, ta_sect, ver, sex_id, urm_id, eop_id, fgn_id, satverbal, satmath, high_sch_gpa, mastered_topics_initial_kc, experiment1, satm_c, satv_c, aleksikc_c, hsgpa_c, final_c, satm_z, satv_z, aleksikc_z, hsgpa_z, final_z)
df_unique_2 <- unique(df_unique_1)
df_unique <- na.omit(df_unique_2)
```
### Make `ta_sect`  unique
```{r}
df_true <- df_unique %>% 
  mutate(ta_sect = ifelse(str_detect(course_fullid, "2016"), paste(ta_sect, "16", sep = "_"), paste(ta_sect, "17", sep = "_")))
```
## Model selection 
### Model selection on `sex_id`
#### Sex Model 1: fixed effect
```{r}
sex_mod1 <- lm(final_c ~ experiment1 * sex_id + 
                 experiment1 + sex_id + 
                 satm_c + satv_c + aleksikc_c + hsgpa_c, 
               data = df_true)
summary(sex_mod1)
```
#### Sex model 2: random effect 
```{r}
sex_mod2 <- lmer(final_c ~ experiment1 * sex_id + 
                   experiment1 + sex_id +
                   satm_c + satv_c + aleksikc_c + hsgpa_c + 
                   (1 | ta_sect),
                 data = df_true, REML = TRUE)
summary(sex_mod2)

AIC(sex_mod1, sex_mod2)
```
* Since the AIC value went up after adding random effect, we can ignore the random effect.  

#### Sex model 3: no interaction or random effect
```{r}
sex_mod3 <- lm(final_c ~ experiment1 + sex_id + 
                 satm_c + satv_c + aleksikc_c + hsgpa_c, 
               data = df_true)
AIC(sex_mod1, sex_mod3)
summary(sex_mod3)
```
* The AIC value decreased for `sex_mod3`
* The interaction between `experiemnt1` and `sex_id` should not be included

#### Sex model 3.a: No `experiment1` 
```{r}
sex_mod3.a <- lm(final_c ~ sex_id + 
                 satm_c + satv_c + aleksikc_c + hsgpa_c, 
               data = df_true)
AIC(sex_mod3, sex_mod3.a)
summary(sex_mod3.a)
```
* The AIC value slightly decreased...
* Since sex_mod3.a is slightly smaller, we keep this model

#### Sex model 3.b: Remove `sex_id`
```{r}
sex_mod3.b <- lm(final_c ~ satm_c + satv_c + aleksikc_c + hsgpa_c, 
               data = df_true)
AIC(sex_mod3.a, sex_mod3.b)
summary(sex_mod3.b)
```
* Since the AIC value increased, `sex_id` should be kept in the model

#### Sex model 3.c: Remove `satv_c`
```{r}
sex_mod3.c <- lm(final_c ~ sex_id + 
                 satm_c + aleksikc_c + hsgpa_c, 
               data = df_true)
AIC(sex_mod3.a, sex_mod3.c)
summary(sex_mod3.c)
```
* The AIC value went up, so `satv_c` should be kept in our model 

#### Sex model 3.d: Remove `aleksikc_c`
```{r}
sex_mod3.d <- lm(final_c ~ sex_id + 
                 satm_c + satv_c + hsgpa_c, 
               data = df_true)
AIC(sex_mod3.a, sex_mod3.d)
summary(sex_mod3.d)
```
* The AIC value went up, `alekikc_c` is kept

#### Sex Model 3.e: remove `hsgpa_c`
```{r}
sex_mod3.e <- lm(final_c ~ sex_id + 
                 satm_c + satv_c + aleksikc_c , 
               data = df_true)
AIC(sex_mod3.a, sex_mod3.e)
summary(sex_mod3.e)
```
* The AIC value went up, `hsgpa_c` is kept

#### Sex Model 3.f: Remove `satm_c`
```{r}
sex_mod3.f <- lm(final_c ~ sex_id + 
                 satv_c + aleksikc_c + hsgpa_c, 
               data = df_true)
AIC(sex_mod3.a, sex_mod3.f)
summary(sex_mod3.f)
```
* Again, the AIC value went up. `satm_c` should be kept
* `sex_mod3.a` is the best fitting model  

### Model selection on `eop_id`
#### Eop model 1: fixed effect
```{r}
eop_mod1 <- lm(final_c ~ experiment1 * eop_id + 
                 experiment1 + eop_id + 
                 satm_c + satv_c + aleksikc_c + hsgpa_c, 
               data = df_true)
summary(eop_mod1)
```
#### EOP model 2: random effect
```{r}
eop_mod2 <- lmer(final_c ~ experiment1 * eop_id + 
                   experiment1 + eop_id +
                   satm_c + satv_c + aleksikc_c + hsgpa_c + 
                   (1 | ta_sect),
                 data = df_true, REML = TRUE)
summary(eop_mod2)
AIC(eop_mod1, eop_mod2)
```
* AIC value went up, random effect should not be removed
* Interaction has the lowest t value, it should be removed next

#### Eop model 3: remove `experiment1` and `eop_id` interaction and no radnom effect 
```{r}
eop_mod3 <- lm(final_c ~ experiment1 + eop_id + 
                 satm_c + satv_c + aleksikc_c + hsgpa_c, 
               data = df_true)
AIC(eop_mod1, eop_mod3)
summary(eop_mod3)
```
* The AIC value decreased with the new model, meaning the interaction should be removed

#### eop model 3.a: remove `eop_id`
```{r}
eop_mod3.a <- lm(final_c ~ experiment1 + 
                 satm_c + satv_c + aleksikc_c + hsgpa_c, 
               data = df_true)
AIC(eop_mod3, eop_mod3.a)
summary(eop_mod3.a)
```
* The AIC value decreased slightly, and the `eop_mod3.a` has the lower value. `eop_id` should be removed

#### eop model 3.b: remove `experiemnt1`
```{r}
eop_mod3.b <- lm(final_c ~ satm_c + satv_c + aleksikc_c + hsgpa_c, 
               data = df_true)
AIC(eop_mod3.a, eop_mod3.b)
summary(eop_mod3.b)
```
* The `eop_mod3.b` has a lower AIC value meaning `experiment1` should also be removed. 

#### eop model 3.c: remove `satv_c`
```{r}
eop_mod3.c <- lm(final_c ~ satm_c + aleksikc_c + hsgpa_c, 
               data = df_true)
AIC(eop_mod3.b, eop_mod3.c)
summary(eop_mod3.c)
```
* The AIC value went up, `satv_c` should be kept. 
* `eop_mod3.c` is the best fitting model





