---
title: "Regression Model"
author: "Ganling"
date: "5/1/2021"
output: 
  html_document:
    keep_md: yes
---
#Load the libraries
```{r}
library(here)
library(tidyverse)
library(skimr)
library(moderndive)
library(ggplot2)
library(dplyr)
library(readr)
```
#Load the data using the directory 
```{r}
proj_dir <- here()
#add original original data directory sting
original_data_dir <- here("original-data", "/")
#add numbers to importable and analysis data directory strings
importable_data_dir <- here("processing-and-analysis", "01-importable-data", "/")
#add analysis directory string
analysis_data_dir <- here("processing-and-analysis", "03-analysis-data", "/")
metadata_dir <- here("original-data", "metadata", "/")
```

Copy data from `original-data` to `01-importable-data`

```{r}
# CFC: Copy rds file from the `original_data_dir` to the `importable_data_dir`.
copy_from <- paste0(original_data_dir, "master_2s_small_deidentified.rds")
copy_to <- paste0(importable_data_dir, "master_2s_small_deidentified.rds")
# CFC: If the file already exists in the target directory, the file will not be 
# overwritten and this command will return `FALSE`.
file.copy(copy_from, copy_to)
```

```{r}
## [1] FALSE
```

Read in data.

```{r}
#CFC: IMport data set from `importable_data_dir`
data_2s <- readRDS("G:/Shared drives/CEG Two-Stage Exams Analysis/Ganling/two-stage-2/processing-and-analysis/01-importable-data/master_2s_small_deidentified.rds")
#data_2s <- readRDS(paste0(importable_data_dir, "master_2s_small_deidenitified.rds")) (still does not work for some reason)
```

#Before starting to make historgrams and boxplots, I would like to remove all the duplicating rows and make a new data frame

```{r}
data_2s_extras_removed = subset(data_2s, select = -c(item_num, stud_ans, exam_key, corr, bloom_rating, complexity_rating_mean, complexity_rating_median, item_code, stem))
data_2s_unique = unique(data_2s_extras_removed, incomparables = FALSE)
View(data_2s_unique)
data_2s_exam_score <- data_2s_unique %>% 
  select(exam1, exam2, finalexam, course_fullid, ver, sex_id, urm_id, eop_id, fgn_id, satverbal, satmath, high_sch_gpa, mastered_topics_initial_kc)
  
data_2s_exam_score_true <- na.omit(data_2s_exam_score)

data_2s_leveled <- data_2s_exam_score_true %>% 
    mutate(course_fullid = factor(course_fullid, 
        levels = c("CHEM_142_A_2016_4", "CHEM_142_B_2016_4", 
                   "CHEM_142_A_2017_4", "CHEM_142_B_2017_4")))
```
#Now, I want to use regression model to see the exam score difference between experiement year and the non experiment year
```{r}
exam1_regression <- data_2s_leveled %>%
  select(course_fullid, exam1)
#now lets find mean and median scores of exam 1
exam1_regression_stats <- exam1_regression %>%
  group_by(course_fullid) %>%
  summarize(exam1_median = median(exam1), 
            exam1_mean = mean(exam1))

#now lets get the regression table for exam 1 
exam1_model <- lm(exam1 ~ course_fullid, data = exam1_regression)
get_regression_table(exam1_model)

#making a data set frame for exam 2 regression
exam2_regression <- data_2s_leveled %>%
  select(course_fullid, exam2)
#Same process for exam 2
exam2_regression_stats <- exam2_regression %>%
  group_by(course_fullid) %>%
  summarize(exam2_median = median(exam2), 
            exam2_mean = mean(exam2))

#now lets get the regression table
exam2_model <- lm(exam2 ~ course_fullid, data = exam2_regression)
get_regression_table(exam2_model)

#making a data set frame for final exam regression
finalexam_regression <- data_2s_leveled %>%
  select(course_fullid, finalexam)
#Same process for exam 2
finalexam_regression_stats <- finalexam_regression %>%
  group_by(course_fullid) %>%
  summarize(finalexam_median = median(finalexam), 
            finalexam_mean = mean(finalexam))

#now lets get the regression table
finalexam_model <- lm(finalexam ~ course_fullid, data = finalexam_regression)
get_regression_table(finalexam_model)
```
#now I would like to see the interaction between the scores when we control for sex
```{r}
exam1_regression_sex <- data_2s_leveled %>%
  select(course_fullid, exam1, sex_id)

#now lets get the regression table for exam 1 control for sex
exam1_sex_model <- lm(exam1 ~ course_fullid * sex_id, data = exam1_regression_sex)
get_regression_table(exam1_sex_model)

exam2_regression_sex <- data_2s_leveled %>%
  select(course_fullid, exam2, sex_id)

#now lets get the regression table for exam 2 control for sex
exam2_sex_model <- lm(exam2 ~ course_fullid * sex_id, data = exam2_regression_sex)
get_regression_table(exam2_sex_model)

finalexam_regression_sex <- data_2s_leveled %>%
  select(course_fullid, finalexam, sex_id)

#now lets get the regression table for final exam control for sex
finalexam_sex_model <- lm(finalexam ~ course_fullid * sex_id, data = finalexam_regression_sex)
get_regression_table(finalexam_sex_model)
```
#EOP status and score interaction
```{r}
exam1_regression_eop <- data_2s_leveled %>%
  select(course_fullid, exam1, eop_id)

#now lets get the regression table for exam 1 control for sex
exam1_eop_model <- lm(exam1 ~ course_fullid * eop_id, data = 
  exam1_regression_eop)
get_regression_table(exam1_eop_model)

exam2_regression_eop <- data_2s_leveled %>%
  select(course_fullid, exam2, eop_id)

#now lets get the regression table for exam 1 control for sex
exam2_eop_model <- lm(exam2 ~ course_fullid * eop_id, data = exam2_regression_eop)
get_regression_table(exam2_eop_model)

finalexam_regression_eop <- data_2s_leveled %>%
  select(course_fullid, finalexam, eop_id)

#now lets get the regression table for exam 1 control for sex
finalexam_eop_model <- lm(finalexam ~ course_fullid * eop_id, data = finalexam_regression_eop)
get_regression_table(finalexam_eop_model)
```
#fgn status and exam score 
```{r}
exam1_regression_fgn <- data_2s_leveled %>%
  select(course_fullid, exam1, fgn_id)

#now lets get the regression table for exam 1 control for sex
exam1_fgn_model <- lm(exam1 ~ course_fullid * fgn_id, data = exam1_regression_fgn)
get_regression_table(exam1_fgn_model)

exam2_regression_fgn <- data_2s_leveled %>%
  select(course_fullid, exam2, fgn_id)

#now lets get the regression table for exam 1 control for sex
exam2_fgn_model <- lm(exam2 ~ course_fullid * fgn_id, data = exam2_regression_fgn)
get_regression_table(exam2_fgn_model)

finalexam_regression_fgn <- data_2s_leveled %>%
  select(course_fullid, finalexam, fgn_id)

#now lets get the regression table for exam 1 control for sex
finalexam_fgn_model <- lm(finalexam ~ course_fullid * fgn_id, data = finalexam_regression_fgn)
get_regression_table(finalexam_fgn_model)
```
#URM status and score interaction
```{r}
exam1_regression_urm <- data_2s_leveled %>%
  select(course_fullid, exam1, urm_id)

#now lets get the regression table for exam 1 control for sex
exam1_urm_model <- lm(exam1 ~ course_fullid * urm_id, data = exam1_regression_urm)
get_regression_table(exam1_urm_model)

exam2_regression_urm <- data_2s_leveled %>%
  select(course_fullid, exam2, urm_id)

#now lets get the regression table for exam 1 control for sex
exam2_urm_model <- lm(exam2 ~ course_fullid * urm_id, data = exam2_regression_urm)
get_regression_table(exam2_urm_model)

finalexam_regression_urm <- data_2s_leveled %>%
  select(course_fullid, finalexam, urm_id)

#now lets get the regression table for exam 1 control for sex
finalexam_urm_model <- lm(finalexam ~ course_fullid * urm_id, data = finalexam_regression_urm)
get_regression_table(finalexam_urm_model)
```
Next, I need to do multiple regression tables for two numerical values 
```{r}
#math SAT scores and exam 1 score 
exam1_regression_satm <- data_2s_leveled %>% 
  select(course_fullid, exam1, satmath)
ggplot(exam1_regression_satm, aes(x = satmath, y = exam1, color = course_fullid)) +
  geom_point() +
  labs(x = "SAT Math Score", y = "Exam 1 score", color = "Class") +
  geom_parallel_slopes(se = FALSE)
```









