---
# title: "Regression Template"
# author: "Colleen Craig"
# date: "8/21/2021"
output: 
  html_document:
    keep_md: yes
---

# Regression Template

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(skimr)
library(knitr)
library(moderndive)
library(janitor) # Tools for exploring and cleaning data
```

## Set Up

### Designate working directories and copy file

```{r}
original_data_dir   <- here("original-data", "/")
importable_data_dir <- here("processing-and-analysis", "01-importable-data", "/")
analysis_data_dir   <- here("processing-and-analysis", "03-analysis-data", "/")
```

Copy the `two-stage-master-wide.rds` file from `original-data-dir` 
to `importable-data-dir` using `file.copy`

```{r}
file.copy()
```

### Load `two-stage-master-wide`

```{r}
read_rds()
```

### Explore the contents of the dtaframe

You won't need all the columns contained here for the current analysis, so 
just select the ones you need. (Consult the [metadata guide](https://github.com/UW-CEG/two-stage-2/blob/master/original-data/metadata/metadata_guide.md) to learn about the
column data.) The df should already contain only unique student records, so you
shouldn't need to remove any duplicates. 

```{r}

```


## Run the model selection procedure using the _CENTERED_ data for all numerical values. 

Follow the guide shown [here](https://docs.google.com/document/d/1XRv2093XdzpSqRVzrrusUUVNMunE6scFvDH-kqAPnPU/edit#). 

Start with the following master model:

`final_c ~ exp * dem_id +` 
          `exp + dem_id +` 
          `satm_c + satv_c + aleksikc_c + hsgpa_c`

### Model selection for `sex_id`

### Model selection for `urm_id`

### Model selection for `eop_id`

### Model selection for `fgn_id`
