---
# title: "En-smallening the Data Set"
# author: "Colleen Craig"
# date: "4/6/2021"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(readxl)
library(janitor)
```

# En-smallening the Data Set
##### Colleen Craig
##### 2021-04-09

## Overview

The first version of the data set that I made in `data_prep.Rmd` is waaaaay too big to be useful 
without access to a computer cluster. Here, I will remake the data set so that it does
not contain per-question information from the two-stage quizzes. For now, I *will* keep the
per-question information from the exams, since we might be able to analyze that in a 
meaningful way.

## Plan of attack:

### Exam, demographics, and prior academic attainment data

I can start with `inst1_df3.rds`, which I created in `data_prep.Rmd`. This file contains
the following variables:

```
"studID", "qtr", "course", "course.fullid", "TAsect", "exam1predict",
'exam2predict', "finalexampredict", "exam1", "exam2", "finalexam", "instructorgrade",
"exam", "ver", "num_items", "length_min", "min_per_ques", "item_num", "stud_ans", 
"key", "corr", "bloom_rating", "complexity_rating_mean", "complexity_rating_median", 
"item_code", "stem", "mastered_topics_initial_kc", "time_initial_kc", "satmath", 
"satverbal", "high_sch_gpa", "cci_total_1", "cci_total_2", "project1", "experiment1", 
"eop.id", "sex.id", "fgn.id", "urm.id", "ethnicity"
```

### Two-stage quiz data

Here, I just want the individual and group scores for each quiz. I will grab the grade 
spreadsheets from Fall 2017 and extract from there. **Should I also extract the exam scores,
and use those in place of the exam scores in `inst1_df3`?**


## Data Import and Manipulation

Set Working Directories

```{r}
proj_dir <- here()
original_data_dir <- here("original-data", "/")
importable_data_dir <- here("processing-and-analysis", "01-importable-data", "/")
analysis_data_dir <- here("processing-and-analysis", "03-analysis-data", "/")
```

Import data:

```{r}
# Load the exams, demographics, and prior academic attainment data 
inst1_df3 <- read_rds(paste0(analysis_data_dir, "inst1_df3.rds"))

# Load the gradebooks from A17 CHEM 142 A & B, which contain the two-stage quiz 
# performance data
chem142_A <- read_excel(paste0(original_data_dir, "CHEM_142A_AUTUMN_2017__--_CRAIG TO UPLOAD - Copy.xls"))
chem142_A
chem142_B <- read_excel(paste0(original_data_dir, "CHEM_142B_AUTUMN_2017_--_CRAIG TO UPLOAD.xls"))
chem142_B
```

### Wrangle A17 Class Data

For `chem142_A` and `chem142_B`, the actual column names are in row 3. I need to pull out
that row from each df and then rename the columns with those values.

```{r}
# Get the row with the actual column names. The output will be a one-row dataframe
A_row3_df <- chem142_A %>% filter(...2 == "Last Name")
B_row3_df <- chem142_B %>% filter(...2 == "Last Name")

# Convert the one-row dataframe to a vector of unnamed character values. NOte that 
# specifying `[1,]` (row 1, all columns) is not strictly necessary because there is
# only one row in these dataframes.
A_colnames <- unname(unlist(A_row3_df[1,]))
str(A_colnames)
B_colnames <- unname(unlist(B_row3_df[1,]))
str(B_colnames)

# Use the base R fxn `colnames` to replace the names of all the columns at once.
colnames(chem142_A) <- A_colnames
colnames(chem142_B) <- B_colnames
```

Note: The original `chem142_B` dataframe has two extra columns compared to the original 
`chem142_A` dataframe, because the lab report scores in A17 142 B had to be normed to reduce
grading discrepancies among TAs. This did not have to be done in A17 142 A. The extra columns
in `chem142_B` are "Lab Rpts, Raw" and "Lab Rpts, Normed".


Remove the rows that don't contain student data.

```{r}
# Remove rows that contain NAs
chem142_A <- chem142_A %>% filter(!is.na(`Last Name`))
chem142_A %>% tail()
chem142_B <- chem142_B %>% filter(!is.na(`Last Name`))
chem142_B %>% tail()

# Remove the new first row, which contains just the column names
chem142_A <- chem142_A %>% filter(Username != "Username")
chem142_B <- chem142_B %>% filter(Username != "Username")

```


Now, select just the columns of interest:

```{r}
chem142_A <- chem142_A %>% select(`Student Number`, `Section/Group`, matches("Q\\d"),
                     "Ex 1 Version", "Ex 1 MC",
                     "Ex 2 Version", "Ex 2 MC Score",
                     `Final Version`, `Final MC Score`)

chem142_B <- chem142_B %>% select(`Student Number`, `Section/Group`, matches("Q\\d"),
                     "Ex 1 Version", "Ex 1 MC",
                     "Ex 2 Version", "Ex 2 MC Score",
                     `Final Version`, `Final MC Score`)

```

### Change Variable Names and Types 

Use the `clean_names` fxn from the `janitor` package to convert all column names to 
"snake case", in which all words are lowercase and "_" is used as a separator. 

```{r}
chem142_A <- chem142_A %>% clean_names()
chem142_B <- chem142_B %>% clean_names()
inst1_df3 <- inst1_df3 %>% clean_names()
```

This introduced an error in the "TA section" column name in `inst1_df3`, so fix that:

```{r}
inst1_df3 <- inst1_df3 %>% rename(ta_sect = t_asect)
```

Convert obvious numbers to numeric type

```{r}
chem142_A <- chem142_A %>% type_convert()
chem142_B <- chem142_B %>% type_convert()
inst1_df3 <- inst1_df3 %>% type_convert()
```

Student ID values in `inst1_df3` remained character type after conversion, but those in
the `chem142_*` dfs were converted to numeric. Not sure why that is. Let's convert
the student IDs back to character in these dataframes, and make them all 7-digits long
as well. I'll have to do some hardcoding cuz I don't have time to figure out how to 
make it generalizable!

```{r}
# Convert student id values to character type
chem142_A$student_number <- chem142_A$student_number %>% as.character()
chem142_B$student_number <- chem142_B$student_number %>% as.character()

# Add leading zeroes to student ids that are less than seven characters. This
# gives them the correct year reference.
chem142_A$student_number[[1]] <- paste0("0", chem142_A$student_number[[1]])
chem142_A$student_number[[2]] <- paste0("0", chem142_A$student_number[[2]])

chem142_B$student_number[[1]] <- paste0("0", chem142_B$student_number[[1]])
chem142_B$student_number[[2]] <- paste0("0", chem142_B$student_number[[2]])
```

For now, I will extract just the quiz data from the `chem142_*` dfs and merge it with 
the `inst1_df3` df. Later I'll create my own quiz + exam dfs from from original gradebooks.

```{r}
chem142_A_quizzes <- chem142_A %>% select(student_number, matches("^q\\d"))
chem142_B_quizzes <- chem142_B %>% select(student_number, matches("^q\\d"))
```

Now row bind the quizzes-only dfs:

```{r}
chem142_quizzes <- bind_rows(chem142_A_quizzes, chem142_B_quizzes)
```

Join `inst1_df3` and `chem142_quizzes`

```{r}
master_2s_small <- left_join(inst1_df3, chem142_quizzes, by = c("stud_id" = "student_number"))
```

Write to disk!

```{r}
write_csv(master_2s_small, paste0(analysis_data_dir, "master_2s_small.csv"))
write_rds(master_2s_small, paste0(analysis_data_dir, "master_2s_small.rds"))
```

#### Deidentify the data for Ganling and Jackson

Create a codebook of student id numbers and two-stage quiz analysis id numbers.

```{r} 
# Extract `studID` from `master_2s_small`
code_book_2s <- master_2s_small %>% distinct(stud_id) 

# Randomize the order of the student numbers
code_book_2s <- code_book_2s %>% sample_frac()

# Add new column to `code_book_2s` for the two-stage id
code_book_2s <- code_book_2s %>% mutate(two_stage_id = NA)

# Create two-stage id values
for (i in 1:nrow(code_book_2s)) {
  code_book_2s$two_stage_id[[i]] <- paste0("chem142_2s_", i) 
}

tail(code_book_2s)

```

Apply code book to `master_two_stage`.

```{r}
# Join `master_2s_small` and `code_book_2s` by `stud_id`
master_2s_small <- master_2s_small %>% left_join(code_book_2s, by = "stud_id") 

# Move `two_stage_id` to the first column 
master_2s_small <- master_2s_small %>% relocate(two_stage_id)

# Remove `stud_id` from `master_2s_small` and save as new df
master_2s_small_deidentified <- master_2s_small %>% select(-stud_id)

```

Write to disk!

```{r}
write_csv(master_2s_small_deidentified, paste0(analysis_data_dir, "master_2s_small_deidentified.csv"))
write_rds(master_2s_small_deidentified, paste0(analysis_data_dir, "master_2s_small_deidentified.rds"))
```