---
title: "Two Stage Analysis Data Prep"
author: "Colleen Craig"
date: "3/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(skimr)
library(moderndive)
library(here)
```
## Set Working Directories

```{r}
proj_dir <- here()
importable_data_dir <- here("processing-and-analysis", "importable-data", "/")
analysis_data_dir <- here("processing-and-analysis", "analysis-data", "/")
metadata_dir <- here("original-data", "metadata", "/")
```

## Data Import

Import data from...

* CHEM 142 exam item analysis: Bloom's and complexity 
* Project Convince: an attempt to find demographics patterns in two-stage quiz project

```{r}
# Leave out `master` for now...I'll need it for CHEM 152 later.
# master <- read_rds(paste0(importable_data_dir, "UW-CHEM-MASTERDATA-A16-A19 (DO NOT TOUCH).rds"))

item_analysis <- read_rds(paste0(importable_data_dir, "UW-CHEM-MASTERDATA-Item-Analysis-A16-A19-20200215 (DO NOT TOUCH).rds"))

metadata <- read_csv(paste0(metadata_dir, "UW-CHEM-MASTERDATA-metadata-20200218.csv"))

convince_0110 <- read_rds(paste0(importable_data_dir, "MASTER-convince-1-10-2019 (DO NOT TOUCH).rds"))
convince_0516 <- read_rds(paste0(importable_data_dir, "MASTER-convince-5-16-19 (DO NOT TOUCH).rds"))

item_analysis %>% distinct(course.fullid)
item_analysis %>% distinct(project1)
```

Looks like I really only need the `item-analysis` df, since it appears to contain 
all the exam item analysis I wrangled in 2019, as well as Michael Mack's previous 
master data.

Actually that's not true...the `item_analysis` df only contains 142 data, since it was 
for the ACTIVATE 142 analysis. I'll have to extract the 152 data from `master` if
I end up pursuing that analysis.

## Data Selection for Two-stage Quiz Analysis in CHEM 142

(_I'll come back to CHEM 152 later..._)

### Filter relevant rows from `item-analysis`

Filter all rows from `item_analysis` that correspond to `inst == inst1` (that's me! =).

```{r}
inst1_df <- item_analysis %>% filter(inst == "inst1")
inst1_df %>% distinct(project1)
inst1_df %>% distinct(experiment1)
inst1_df %>% distinct(course.fullid, experiment1, project1)
```
Some of the 2016 data have `NA` instead of `CONTROL` for `experiment1`, and some 
of the 2017 data have `NA` instead of `EXPERIMENTAL` for `experiment1`. I'll explore this
after winnowing down the columns to just those I care about.

Now, select the following columns: `studID`, `qtr`, `course`, `course.fullid`, `TAsect`, `exam1predict`,
`exam2predict`, `finalexampredict`, `exam1`, `exam2`, `finalexam`, `instructorgrade`,
`exam`, `ver`, `num_items`, `length_min`, `min_per_ques`, `item_num`, `stud_ans`, `key`, `corr`, `bloom_rating`, `complexity_rating_mean`, `complexity_rating_median`, `item_code`, `stem`, `mastered_topics_initial_kc`, `time_initial_kc`, `satmath`, `satverbal`, `high_sch_gpa`, `cci_total_1`, `cci_total_2`, `project1`, `experiment1`, `eop.id`, `sex.id`, `fgn.id`, `urm.id`, `ethnicity` 

```{r}
inst1_df2 <- inst1_df %>% 
  select("studID", "qtr", "course", "course.fullid", "TAsect", "exam1predict",
         'exam2predict', "finalexampredict", "exam1", "exam2", "finalexam", "instructorgrade",
         "exam", "ver", "num_items", "length_min", "min_per_ques", "item_num", "stud_ans", 
         "key", "corr", "bloom_rating", "complexity_rating_mean", "complexity_rating_median", 
         "item_code", "stem", "mastered_topics_initial_kc", "time_initial_kc", "satmath", 
         "satverbal", "high_sch_gpa", "cci_total_1", "cci_total_2", "project1", "experiment1", 
         "eop.id", "sex.id", "fgn.id", "urm.id", "ethnicity")

inst1_df2 %>% distinct(qtr, course.fullid, project1) %>% arrange(project1)
```

Finally, `filter` out just the 2016 and 2017 data:

```{r}
inst1_df3 <- inst1_df2 %>% filter((qtr == "A16") | (qtr == "A17"))

inst1_df3 %>% distinct(qtr, course.fullid, project1) %>% arrange(project1)
```


```{r}
#inst1_df2 %>% filter(is.na(experiment1))
na_exp1 <- inst1_df3 %>% filter(is.na(experiment1)) %>% distinct(studID) 
na_proj1 <- inst1_df3 %>% filter(is.na(project1)) %>% distinct(studID) 

na_exp1 == na_proj1
```

The same students that have `NA` for `experiment1` also have `NA` for `project1`.

```{r}
inst1_df3 %>% select(studID, course.fullid, satmath) %>% 
  filter(is.na(satmath)) %>% 
  distinct(studID) %>% 
  tally()
```
There are 95 students who have `NA` for `satamth`. 

I will also have to track down the two-stage quiz data and combine it with the `inst1_df3`
data.

Save `inst1_df3` to the analysis-data folder:

```{r}
write_rds(inst1_df3, paste0(analysis_data_dir, "inst1_df3.rds"))
```

### Explore the two-stage quiz datasets created in Project Convince

From the Project Convince exploration in 2019, there were two master data files created:

* MASTER-convince-1-10-2019 (DO NOT TOUCH).rds
* MASTER-convince-5-16-19 (DO NOT TOUCH).rds

I found some scripts in the `ChEdR Project Convince` share drive that were used to 
create at least the Jan 10 master file. I put copies of them in a subfolder of `command-files`.
`They were written `Filter.R` and `First-response.R` were written by Michael Mack, 
and `correct.R` was written by Abbie Pickering, based on `First-response.R`. I have not
dug too much into the code in these scripts yet, but `Filter.R` definitely creates the Jan 10
master data set (csv and rds versions), and `First-response.R` conducts some regression
analyses on the two-stage data. (This file appears to have been created in Dec 2018, 
but I remember Michael telling me about the disappointing lack of a signal in the 
two-stage analysis long before then. I don't (currently) know where that analysis is.
The one here in `First-response.R` must be his first noodlings on Project Convince
before we brought Abbie and Sam Gunnerson on board.)

There is no reference to the creation of the May 16 master data set in any of these 
scripts, though it must have been created in R since it is an rds file. 

ANYWAY. Let's take a look at these two data sets to see what the difference is.

```{r}
head(convince_0110)
head(convince_0516)
```
Differences between the 01-10 and 05-16 master data sets:

* These dfs have the same number of rows (82,910), but the 0516 df has one more column 
(444 vs. 443). 
* The extra column in 0516 is the very first one, simply named `X`, containing integers.
Generally the integers are increasing from top to bottom and are the same as the row numbers,
except occasionally there is a very different number...see row 5 in 0516, where `X = 5733`.
But, row 5733 in 0110 has the same data as row 5 in 0516, except for the `X` column, and
trivial differences in the date formatin the `final_kc_start_date` and `final_kc_end_date`
columns. (Check out the following code chunk for a brute force comparison of these rows.)
* There are also some cases in the 0110 dataset in which the `group.id` has an underscore,
and these are removed in the 0516 dataset.

```{r}
jan10 <- convince_0110 %>% slice(5733)
may16 <- convince_0516 %>% slice(5)

may16 <- may16 %>% select(-X)

combo <- bind_rows(jan10, may16)

```

I think the `X` column is related to some kind of error checking that Sam Gunnerson
did, but in the code she [uploaded to Github](https://github.com/cfcraig/project-convince),
I can't find a reference to creating the 0516 file anywhere, although it is the one she 
opens in [this script](https://github.com/cfcraig/project-convince/blob/master/Sam's%20Scripts/useful/1_condense_build.R)
(see line 47). 

It's also not clear why certain rows were reordered, in ways that 
don't make sense. Row 5 in 0516 contains data for a completely different student from 
a completely different sewction 
than in rows 1-4 and row 6 (and beyond). There seem to be other rows like it. Further,
since the data in these two rows is identical, it doesn't appear that there were any 
corrections made when 0110 was converted to 0516. 

(_I need to dig into this mystery further...consult the Project Convince meeting notes..._)

FOR NOW: I will push forward with the 0516 dataset. 

### Extract relevant data from the 05-16 master data set

```{r}
convince_0516
```
Now, select the following columns from 0516 dataset: `studentnum`, `course.fullid`, `group.id`,
`questionnum`, `quiznum`, `all.response`, `first.response`, `responsenum`, `ind.response`,
`quizversion`, `key`, `project1`

```{r}
two_stage <- convince_0516 %>% 
  select("studentnum", "course.fullid", "group.id", "questionnum", "quiznum", 
         "all.response", "first.response", "responsenum", "ind.response", "quizversion", 
         "key", "project1")

two_stage %>% distinct(course.fullid, project1) %>% arrange(project1)
two_stage %>% distinct(course.fullid, project1)
two_stage %>% distinct(course.fullid) 
```

### Combine the data sets 

Since two-stage quiz data are only in 2017_4, and there may be students who took 142 
in both 2016 and 2017, I will create a 2017_4-only df from `inst1_df3` and then 
join it to the `two-stage` df by student id. Then, I'll use `bind_rows` to combine the 
2017_4 data with 2016_4.

```{r}
# Both dfs have a `key` variable, so update names to make them more specific
inst1_df3 <- inst1_df3 %>% rename(exam_key = key)
two_stage <- two_stage %>% rename(quiz_key = key)
```


#### First, create year-specific versions of the `inst1_df3` data

Filter out the 2017_4 data.

```{r}
only_2017_4 <- 
  inst1_df3 %>% filter(str_detect(course.fullid, "2017_4"))

only_2017_4 %>% distinct(course.fullid)
```

Filter out everything else.

```{r}
not_2017_4 <- 
  inst1_df3 %>% filter(!str_detect(course.fullid, "2017_4"))

not_2017_4

not_2017_4 %>% distinct(course.fullid)
```

#### Next, `left_join` the `only_2017_4` and `two_stage` dfs

```{r}
# Change `studentnum` in `two_stage` from integer to character
two_stage$studentnum <- as.character(two_stage$studentnum)

# Remove `project1` and `course.fullid` columns from `two_stage` since they are 
# redundant with `only_2017_4`
two_stage <- two_stage %>% select(!project1)
two_stage <- two_stage %>% select(!course.fullid)

joined_2017_4 <- 
  left_join(only_2017_4, two_stage, by = c("studID" = "studentnum"))
```

#### Finally, `bind_row`s between `not_2017_4` and `joined_2017_4`

```{r}
master_two_stage <- 
  bind_rows(joined_2017_4, not_2017_4)
```


#### Deidentify the data for Ganling and Jackson

Create a codebook of student id numbers and two-stage quiz analysis id numbers.

```{r} 
# Extract `studID` from `master_two_stage`
code_book_2s <- master_two_stage %>% distinct(studID) 

# Randomize the order of the student numbers
code_book_2s <- code_book_2s %>% sample_frac()

# Add new column to `code_book_2s` for the two-stage id
code_book_2s <- code_book_2s %>% mutate(two_stage_id = NA)

# Create two-stage id values
for (i in 1:nrow(code_book_2s)) {
  code_book_2s$two_stage_id[[i]] <- paste0("chem142_2s_", i) 
}

tail(code_book_2s)

```

Apply code book to `master_two_stage`.

```{r}
# Join `master_two_stage` and `code_book_2s` by `studID`
master_two_stage <- master_two_stage %>% left_join(code_book_2s, by = "studID") 

# Move `two_stage_id` to the first column 
master_two_stage <- master_two_stage %>% relocate(two_stage_id)

# Remove `studID` from `master_two_stage` and save as new df
master_2s_deidentified <- master_two_stage %>% select(-studID)

```

Save to disk!

```{r}
write_rds(master_two_stage, paste0(analysis_data_dir, "master_two_stage.rds"))
write_csv(master_two_stage, paste0(analysis_data_dir, "master_two_stage.csv"))
write_rds(master_2s_deidentified, paste0(analysis_data_dir, "master_2s_deidentified.rds"))
write_csv(master_2s_deidentified, paste0(analysis_data_dir, "master_2s_deidentified.csv"))
```

