---
title: 'Two Stage Data Visualization'
author: "Jackson Hughes"
date: "1/3/2022"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(skimr)
library(knitr)
library(moderndive)
library(janitor)
library(lme4)
```

## Set up

### Designate working directories and copy file

```{r}
proj_dir <- here()
original_data_dir   <- here("original-data", "/")
importable_data_dir <- here("processing-and-analysis", "01-importable-data", "/")
analysis_data_dir   <- here("processing-and-analysis", "03-analysis-data", "/")
metadata_dir <- here("original-data", "metadata", "/")

copy_from <- paste0(original_data_dir, "two_stage_master_wide_deid.rds")
copy_to <- paste0(importable_data_dir, "two_stage_master_wide_deid.rds")
file.copy(copy_from, copy_to)
```

### Import dataset

```{r}
master_original_1 <- readRDS(paste0(importable_data_dir, "two_stage_master_wide_deid.rds"))
```

### Change value of NAs in `exp` column to `EXPERIMENT`

```{r}
master_original_2 <- master_original_1 %>% 
  mutate(exp = replace_na(exp, "EXPERIMENT"))
```


### Select relevant columns

```{r}
master <- master_original_2 %>% 
  select(two_stage_id, course_fullid, exp, exam1, exam2, ta_sect, sex_id, urm_id, eop_id, fgn_id, satm, satv, aleksikc_score, hs_gpa, final)
```

## Data Visualization

### Box Blots

#### Final exam score vs. `sex_id`

```{r}
ggplot(data = master, mapping = aes(x = sex_id, y = final)) +
  geom_boxplot()
```

#### Final exam score vs. `urm_id`

```{r}
ggplot(data = master, mapping = aes(x = urm_id, y = final)) +
  geom_boxplot()
```

#### Final exam score vs. `eop_id`

```{r}
ggplot(data = master, mapping = aes(x = eop_id, y = final)) +
  geom_boxplot()
```

#### Final exam score vs. `fgn_id`

```{r}
ggplot(data = master, mapping = aes(x = fgn_id, y = final)) +
  geom_boxplot()
```

#### `satm` vs. `sex_id`

```{r}
ggplot(data = master, mapping = aes(x = sex_id, y = satm)) +
  geom_boxplot()
```

#### `satm` vs. `urm_id`

```{r}
ggplot(data = master, mapping = aes(x = urm_id, y = satm)) +
  geom_boxplot()
```

#### `satm` vs. `eop_id`

```{r}
ggplot(data = master, mapping = aes(x = eop_id, y = satm)) +
  geom_boxplot()
```

#### `satm` vs. `fgn_id`

```{r}
ggplot(data = master, mapping = aes(x = fgn_id, y = satm)) +
  geom_boxplot()
```

#### `aleksikc_score` vs. `sex_id`

```{r}
ggplot(data = master, mapping = aes(x = sex_id, y = aleksikc_score)) +
  geom_boxplot()
```

#### `aleksikc_score` vs. `urm_id`

```{r}
ggplot(data = master, mapping = aes(x = urm_id, y = aleksikc_score)) +
  geom_boxplot()
```

#### `aleksikc_score` vs. `eop_id`

```{r}
ggplot(data = master, mapping = aes(x = eop_id, y = aleksikc_score)) +
  geom_boxplot()
```

#### `aleksikc_score` vs. `fgn_id`

```{r}
ggplot(data = master, mapping = aes(x = fgn_id, y = aleksikc_score)) +
  geom_boxplot()
```

#### `hs_gpa` vs. `sex_id`

```{r}
ggplot(data = master, mapping = aes(x = sex_id, y = hs_gpa)) +
  geom_boxplot()
```

#### `hs_gpa` vs. `urm_id`

```{r}
ggplot(data = master, mapping = aes(x = urm_id, y = hs_gpa)) +
  geom_boxplot()
```

#### `hs_gpa` vs. `eop_id`

```{r}
ggplot(data = master, mapping = aes(x = eop_id, y = hs_gpa)) +
  geom_boxplot()
```

#### `hs_gpa` vs. `fgn_id`

```{r}
ggplot(data = master, mapping = aes(x = fgn_id, y = hs_gpa)) +
  geom_boxplot()
```

### Scatterplots

#### Create subsets of original data for each demographic ID

```{r}
master_male <- master %>% 
  filter(sex_id == "Male")
master_female <- master %>% 
  filter(sex_id == "Female")
master_urm <- master %>% 
  filter(urm_id == "URM")
master_non_urm <- master %>% 
  filter(urm_id == "non-URM")
master_eop <- master %>% 
  filter(eop_id == "EOP")
master_non_eop <- master %>% 
  filter(eop_id == "non-EOP")
master_fgn <- master %>% 
  filter(fgn_id == "FGN")
master_non_fgn <- master %>% 
  filter(fgn_id == "non-FGN")
```

#### Final exam score vs. `satm` (grouped by `sex_id`)

```{r}
ggplot() +
  geom_point(data = master_male, aes(x = satm, y = final), color = "blue") +
  geom_smooth(data = master_male, aes(x = satm, y = final), method = 'lm', formula = y ~ x, color = "blue") +
  geom_point(data = master_female, aes(x = satm, y = final), color = "red") +
  geom_smooth(data = master_female, aes(x = satm, y = final), method = 'lm', formula = y ~ x, color = "red")
```

#### Final exam score vs. `aleksikc_score` (grouped by `sex_id`)

```{r}
ggplot() +
  geom_point(data = master_male, aes(x = aleksikc_score, y = final), color = "blue") +
  geom_smooth(data = master_male, aes(x = aleksikc_score, y = final), method = 'lm', formula = y ~ x, color = "blue") +
  geom_point(data = master_female, aes(x = aleksikc_score, y = final), color = "red") +
  geom_smooth(data = master_female, aes(x = aleksikc_score, y = final), method = 'lm', formula = y ~ x, color = "red")
```

#### Final exam score vs. `hs_gpa` (grouped by `sex_id`)

```{r}
ggplot() +
  geom_point(data = master_male, aes(x = hs_gpa, y = final), color = "blue") +
  geom_smooth(data = master_male, aes(x = hs_gpa, y = final), method = 'lm', formula = y ~ x, color = "blue") +
  geom_point(data = master_female, aes(x = hs_gpa, y = final), color = "red") +
  geom_smooth(data = master_female, aes(x = hs_gpa, y = final), method = 'lm', formula = y ~ x, color = "red")
```

#### Final exam score vs. `satm` (grouped by `urm_id`)

```{r}
ggplot() +
  geom_point(data = master_urm, aes(x = satm, y = final), color = "red") +
  geom_smooth(data = master_urm, aes(x = satm, y = final), method = 'lm', formula = y ~ x, color = "red") +
  geom_point(data = master_non_urm, aes(x = satm, y = final), color = "black") +
  geom_smooth(data = master_non_urm, aes(x = satm, y = final), method = 'lm', formula = y ~ x, color = "black")
```

#### Final exam score vs. `aleksikc_score` (grouped by `urm_id`)

```{r}
ggplot() +
  geom_point(data = master_urm, aes(x = aleksikc_score, y = final), color = "red") +
  geom_smooth(data = master_urm, aes(x = aleksikc_score, y = final), method = 'lm', formula = y ~ x, color = "red") +
  geom_point(data = master_non_urm, aes(x = aleksikc_score, y = final), color = "black") +
  geom_smooth(data = master_non_urm, aes(x = aleksikc_score, y = final), method = 'lm', formula = y ~ x, color = "black")
```

#### Final exam score vs. `hs_gpa` (grouped by `urm_id`)

```{r}
ggplot() +
  geom_point(data = master_urm, aes(x = hs_gpa, y = final), color = "red") +
  geom_smooth(data = master_urm, aes(x = hs_gpa, y = final), method = 'lm', formula = y ~ x, color = "red") +
  geom_point(data = master_non_urm, aes(x = hs_gpa, y = final), color = "black") +
  geom_smooth(data = master_non_urm, aes(x = hs_gpa, y = final), method = 'lm', formula = y ~ x, color = "black")
```

#### Final exam score vs. `satm` (grouped by `eop_id`)

```{r}
ggplot() +
  geom_point(data = master_eop, aes(x = satm, y = final), color = "red") +
  geom_smooth(data = master_eop, aes(x = satm, y = final), method = 'lm', formula = y ~ x, color = "red") +
  geom_point(data = master_non_eop, aes(x = satm, y = final), color = "black") +
  geom_smooth(data = master_non_eop, aes(x = satm, y = final), method = 'lm', formula = y ~ x, color = "black")
```

#### Final exam score vs. `aleksikc_score` (grouped by `eop_id`)

```{r}
ggplot() +
  geom_point(data = master_eop, aes(x = aleksikc_score, y = final), color = "red") +
  geom_smooth(data = master_eop, aes(x = aleksikc_score, y = final), method = 'lm', formula = y ~ x, color = "red") +
  geom_point(data = master_non_eop, aes(x = aleksikc_score, y = final), color = "black") +
  geom_smooth(data = master_non_eop, aes(x = aleksikc_score, y = final), method = 'lm', formula = y ~ x, color = "black")
```

#### Final exam score vs. `hs_gpa` (grouped by `eop_id`)

```{r}
ggplot() +
  geom_point(data = master_eop, aes(x = hs_gpa, y = final), color = "red") +
  geom_smooth(data = master_eop, aes(x = hs_gpa, y = final), method = 'lm', formula = y ~ x, color = "red") +
  geom_point(data = master_non_eop, aes(x = hs_gpa, y = final), color = "black") +
  geom_smooth(data = master_non_eop, aes(x = hs_gpa, y = final), method = 'lm', formula = y ~ x, color = "black")
```

#### Final exam score vs. `satm` (grouped by `fgn_id`)

```{r}
ggplot() +
  geom_point(data = master_fgn, aes(x = satm, y = final), color = "red") +
  geom_smooth(data = master_fgn, aes(x = satm, y = final), method = 'lm', formula = y ~ x, color = "red") +
  geom_point(data = master_non_fgn, aes(x = satm, y = final), color = "black") +
  geom_smooth(data = master_non_fgn, aes(x = satm, y = final), method = 'lm', formula = y ~ x, color = "black")
```

#### Final exam score vs. `aleksikc_score` (grouped by `fgn_id`)

```{r}
ggplot() +
  geom_point(data = master_fgn, aes(x = aleksikc_score, y = final), color = "red") +
  geom_smooth(data = master_fgn, aes(x = aleksikc_score, y = final), method = 'lm', formula = y ~ x, color = "red") +
  geom_point(data = master_non_fgn, aes(x = aleksikc_score, y = final), color = "black") +
  geom_smooth(data = master_non_fgn, aes(x = aleksikc_score, y = final), method = 'lm', formula = y ~ x, color = "black")
```

#### Final exam score vs. `hs_gpa` (grouped by `fgn_id`)

```{r}
ggplot() +
  geom_point(data = master_fgn, aes(x = hs_gpa, y = final), color = "red") +
  geom_smooth(data = master_fgn, aes(x = hs_gpa, y = final), method = 'lm', formula = y ~ x, color = "red") +
  geom_point(data = master_non_fgn, aes(x = hs_gpa, y = final), color = "black") +
  geom_smooth(data = master_non_fgn, aes(x = hs_gpa, y = final), method = 'lm', formula = y ~ x, color = "black")
```

### Boxplots

#### `sex_id` distribution between control and experimental years

```{r}
ggplot(data = master, aes(x = sex_id)) +
  geom_bar() +
  facet_wrap(~ exp)
```

#### `urm_id` distribution between control and experimental years

```{r}
ggplot(data = master, aes(x = urm_id)) +
  geom_bar() +
  facet_wrap(~ exp)
```

#### `eop_id` distribution between control and experimental years

```{r}
ggplot(data = master, aes(x = eop_id)) +
  geom_bar() +
  facet_wrap(~ exp)
```

#### `fgn_id` distribution between control and experimental years

```{r}
ggplot(data = master, aes(x = fgn_id)) +
  geom_bar() +
  facet_wrap(~ exp)
```