---
title: 'MLM With Unique TA Sections'
author: "Jackson Hughes"
date: "6/28/2021"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(skimr)
library(moderndive)
library(lme4)
```

# Preliminary setup and data wrangling

## Data import

### Set working directories

```{r}
proj_dir <- here()
original_data_dir <- here("original-data", "/")
importable_data_dir <- here("processing-and-analysis", "01-importable-data", "/")
analysis_data_dir <- here("processing-and-analysis", "03-analysis-data", "/")
metadata_dir <- here("original-data", "metadata", "/")
```

### Copy data from `original-data` to `01-importable-data`

```{r}
copy_from <- paste0(original_data_dir, "master_2s_small_deid_scaled.rds")
copy_to <- paste0(importable_data_dir, "master_2s_small_deid_scaled.rds")
file.copy(copy_from, copy_to)
```

### Import master dataset

```{r}
master <- readRDS(paste0(importable_data_dir, "master_2s_small_deid_scaled.rds"))
```

## Data wrangling

### Condense dataset to only include unique rows; omit NAs

```{r}
master_extras_removed <- subset(master, select = -c(item_num, stud_ans, exam_key, corr, bloom_rating, complexity_rating_mean, complexity_rating_median, item_code, stem))
master_unique_1 <- unique(master_extras_removed, incomparables = FALSE)
master_unique_2 <- master_unique_1 %>% 
  select(exam1, exam2, finalexam, course_fullid, ta_sect, ver, sex_id, urm_id, eop_id, fgn_id, satverbal, satmath, high_sch_gpa, mastered_topics_initial_kc, experiment1, satm_c, satv_c, aleksikc_c, hsgpa_c, final_c, satm_z, satv_z, aleksikc_z, hsgpa_z, final_z)
master_unique <- na.omit(master_unique_2)
```

### Make `ta_sect` values unique across `course_fullid`

#### I'm going to be trying the code that CFC showed us

```{r, eval = FALSE}
master_unique_3 <- master_unique %>% 
  select(course_fullid)
master_unique_4 <- master_unique_3 %>% 
  mutate(ta_sect = str_extract(course_fullid, "\\d\\d\\d\\d")) "\\d{4}"
#### This line ^^ is the one that is giving me trouble
master_true <- master_unique_3 %>% 
  mutate(ta_sect = ifelse(str_detect(course_fullid, "2016"), paste(ta_sect, "16", sep = "_"), paste(ta_sect, "17", sep = "_")))
```

#### That isn't working... for now I'm just going to split the dataframe and use `paste0`.

```{r}
master_unique$ta_sect <- as.character(master_unique$ta_sect)

master_unique_2016 <- subset(master_unique, course_fullid != "CHEM_142_A_2017_4" & course_fullid != "CHEM_142_B_2017_4")
master_unique_2017 <- subset(master_unique, course_fullid != "CHEM_142_A_2016_4" & course_fullid != "CHEM_142_B_2016_4")

master_unique_2016$ta_sect <- paste0(master_unique_2016$ta_sect, "_16")
master_unique_2017$ta_sect <- paste0(master_unique_2017$ta_sect, "_17")

master_true <- rbind(master_unique_2016, master_unique_2017)
master_true$ta_sect <- as.factor(master_true$ta_sect)
```

# Regression

## Regression of `final_c` on `experiment1` with control variables (no MLM)

```{r}
mod_1 <- lm(final_c ~ experiment1 + satm_c + satv_c + aleksikc_c + hsgpa_c, data = master_true)
summary(mod_1)
```

## MLM random intercepts model for TA section (centered data)

```{r}
mod_2 <- lmer(final_c ~ experiment1 + satm_c + satv_c + aleksikc_c + hsgpa_c + (1|ta_sect), data = master_true)
summary(mod_2)
```

## Random effects models including demographic interactions

### Interactions from `sex_id`

```{r}
sex_id_mod <- lmer(final_c ~ experiment1*sex_id + satm_c + satv_c + aleksikc_c + hsgpa_c + (1|ta_sect), data = master_true)
summary(sex_id_mod)
```

### Interactions from `urm_id`

```{r}
urm_id_mod <- lmer(final_c ~ experiment1*urm_id + satm_c + satv_c + aleksikc_c + hsgpa_c + (1|ta_sect), data = master_true)
summary(urm_id_mod)
```

### Interactions from `eop_id`

```{r}
eop_id_mod <- lmer(final_c ~ experiment1*eop_id + satm_c + satv_c + aleksikc_c + hsgpa_c + (1|ta_sect), data = master_true)
summary(eop_id_mod)
```

### Interactions from `fgn_id`

```{r}
fgn_id_mod <- lmer(final_c ~ experiment1*fgn_id + satm_c + satv_c + aleksikc_c + hsgpa_c + (1|ta_sect), data = master_true)
summary(fgn_id_mod)
```

## Model Selection Using AIC

#### Standard regression model vs. random intercepts model (no demographic interactions)

```{r}
AIC(mod_1, mod_2)
```

#### AIC comparison of all of the demographic interactions random effects models

```{r}
AIC(sex_id_mod, urm_id_mod, eop_id_mod, fgn_id_mod)
```

## Notes
* I'd like to discuss regex some more... I had some trouble implementing the CFC code from the google doc into this data set.