---
title: 'Model Selection with the New Dataset'
author: "Jackson Hughes"
date: "8/30/2021"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(skimr)
library(knitr)
library(moderndive)
library(janitor)
library(lme4)
```

## Set up

### Designate working directories and copy file

```{r}
proj_dir <- here()
original_data_dir   <- here("original-data", "/")
importable_data_dir <- here("processing-and-analysis", "01-importable-data", "/")
analysis_data_dir   <- here("processing-and-analysis", "03-analysis-data", "/")
metadata_dir <- here("original-data", "metadata", "/")

copy_from <- paste0(original_data_dir, "two_stage_master_wide_deid.rds")
copy_to <- paste0(importable_data_dir, "two_stage_master_wide_deid.rds")
file.copy(copy_from, copy_to)
```

### Import dataset

```{r}
master_original_1 <- readRDS(paste0(importable_data_dir, "two_stage_master_wide_deid.rds"))
```

### Change value of NAs in `exp` column to `EXPERIMENT`

```{r}
master_original_2 <- master_original_1 %>% 
  mutate(exp = replace_na(exp, "EXPERIMENT"))
```


### Select relevant columns

```{r}
master <- master_original_2 %>% 
  select(two_stage_id, course_fullid, exp, exam1, exam2, ta_sect, sex_id, urm_id, eop_id, fgn_id, satm_c, satv_c, aleksikc_c, hs_gpa_c, final_c, satm_z, satv_z, aleksikc_z, hs_gpa_z, final_z)
```

# Model Selection

## Model selection for `sex_id`

### 1. Fixed-effects only model (sex)

```{r}
sex_mod1 <- lm(final_c ~ exp * sex_id +
                         exp + sex_id +
                         satm_c + satv_c + aleksikc_c + hs_gpa_c,
                         data = master)
summary(sex_mod1)
```

### 2. Random intercepts model with `ta_sect`; REML = TRUE (sex)

```{r}
sex_mod2 <- lmer(final_c ~ exp * sex_id +
                         exp + sex_id +
                         satm_c + satv_c + aleksikc_c + hs_gpa_c +
                         (1 | ta_sect),
                         data = master, REML = TRUE)
summary(sex_mod2)
```

#### AIC comparison of `sex_mod1` and `sex_mod2`

```{r}
AIC(sex_mod1, sex_mod2)
```

* The AIC *decreases* by more than 2 for `sex_mod2`, so random effects are not retained in the model.

### 3. Fixed-effects only model with *no* demographic interactions (sex)

```{r}
sex_mod1.a <- lm(final_c ~ exp + sex_id +
                           satm_c + satv_c + aleksikc_c + hs_gpa_c,
                           data = master)
summary(sex_mod1)
```

#### AIC comparison of `sex_mod1` and `sex_mod1.a`

```{r}
AIC(sex_mod1, sex_mod1.a)
```

* The AIC *decreases* by *less than 2* for `sex_mod1.a`, so we will move forward with the simplest model, which is `sex_mod1.a`. Therefore, demographic interactions will not be retained in this model.

### 4. Removal of parameters one by one

* We will start with removing `exp`, which has the lowest t-value

```{r}
sex_mod1.b <- lm(final_c ~ sex_id +
                           satm_c + satv_c + aleksikc_c + hs_gpa_c,
                           data = master)
summary(sex_mod1.b)
```

#### AIC comparison of `sex_mod1.a` and `sex_mod1.b`

```{r}
AIC(sex_mod1.a, sex_mod1.b)
```

* The AIC *decreases* by *less than 2* for `sex_mod1.b`, so we will retain this model since it is the simplest. 

* Now we will try removing `sex_id`

```{r}
sex_mod1.c <- lm(final_c ~ satm_c + satv_c + aleksikc_c + hs_gpa_c,
                           data = master)
summary(sex_mod1.c)
```

#### AIC comparison of `sex_mod1.b` and `sex_mod1.c`

```{r}
AIC(sex_mod1.b, sex_mod1.c)
```

* The AIC *increases* by *more than 2* for `sex_mod1.c`, so we will retain the `sex_id` variable in our model.
* Our final model is `sex_mod1.b`, which is summarized above.

## Model selection for `urm_id`

### 1. Fixed-effects only model (URM)

```{r}
urm_mod1 <- lm(final_c ~ exp * urm_id +
                         exp + urm_id +
                         satm_c + satv_c + aleksikc_c + hs_gpa_c,
                         data = master)
summary(urm_mod1)
```

### 2. Random intercepts model with `ta_sect`; REML = TRUE (URM)

```{r}
urm_mod2 <- lmer(final_c ~ exp * urm_id +
                         exp + urm_id +
                         satm_c + satv_c + aleksikc_c + hs_gpa_c +
                         (1 | ta_sect),
                         data = master, REML = TRUE)
summary(urm_mod2)
```

#### AIC comparison of `urm_mod1` and `urm_mod2`

```{r}
AIC(urm_mod1, urm_mod2)
```

* The AIC *increases* by *more than 2* for `urm_mod2`, so we will not include random effects from `ta_sect` in our model.

### 3. Fixed-effects only model with *no* demographic interactions (URM)

```{r}
urm_mod1.a <- lm(final_c ~ exp + urm_id +
                           satm_c + satv_c + aleksikc_c + hs_gpa_c,
                           data = master)
summary(urm_mod1.a)
```

#### AIC comparison of `urm_mod1` and `urm_mod1.a`

```{r}
AIC(urm_mod1, urm_mod1.a)
```

* The AIC *decreases* by *less than 2* for `urm_mod1.a`, so we will retain this model because it is the simplest.

### 4. Removal of parameters one by one

* Start by removing `urm_id` because it has the lowest t-value

```{r}
urm_mod1.b <- lm(final_c ~ exp +
                           satm_c + satv_c + aleksikc_c + hs_gpa_c,
                           data = master)
summary(urm_mod1.b)
```

#### AIC comparison of `urm_mod1.a` and `urm_mod1.b`

```{r}
AIC(urm_mod1.a, urm_mod1.b)
```

* The AIC *increases* by *more than 2* for `urm_mod1.b`, so we will retain `urm_id` in our model.

* Now let's try removing `exp` from our model:

```{r}
urm_mod1.c <- lm(final_c ~ urm_id +
                           satm_c + satv_c + aleksikc_c + hs_gpa_c,
                           data = master)
summary(urm_mod1.c)
```

#### AIC comparison of `urm_mod1.a` and `urm_mod1.c`

```{r}
AIC(urm_mod1.a, urm_mod1.c)
```

* The AIC *decreases* by *less than 2* for `urm_mod1.c`, which is the simplest model, so we will remove the `exp` parameter from our model.

* Now let's try removing `satv_c`

```{r}
urm_mod1.d <- lm(final_c ~ urm_id +
                           satm_c + aleksikc_c + hs_gpa_c,
                           data = master)
summary(urm_mod1.d)
```

#### AIC comparison of `urm_mod1.c` and `urm_mod1.d`

```{r}
AIC(urm_mod1.c, urm_mod1.d)
```

* The AIC *increases* by *more than 2* for `urm_mod1.d`, so we will retain the `satv_c` parameter in our model.
* Our final model is `urm_mod1.c`, which is summarized above.