---
title: "Exploring POLR with Survey Data"
author: "Jackson Hughes"
date: "2/14/2022"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(skimr)
library(moderndive)
library(lme4)
library(MASS) # fxns and datasets to support "Modern Applied Statistics with S"
library(effects)
library(nnet) # for neural networks and Multinomial Log-Linear Models
library(Hmisc) # misc fxns for data analysis
library(car)
```

## Set up

### Designate working directories and copy file

```{r}
original_data_dir   <- here("original-data")
importable_data_dir <- here("processing-and-analysis", "01-importable-data")
analysis_data_dir   <- here("processing-and-analysis", "03-analysis-data")
metadata_dir <- here("original-data", "metadata")
copy_from <- here(original_data_dir, "two_stage_master_wide_deid.rds")
copy_to <- here(importable_data_dir, "two_stage_master_wide_deid.rds")
file.copy(copy_from, copy_to)
```

### Import dataset

```{r}
master_original_1 <- readRDS(here(importable_data_dir, "two_stage_master_wide_deid.rds"))
```

### Change value of NAs in `exp` column to `EXPERIMENT`

```{r}
master <- master_original_1 %>% 
  mutate(exp = replace_na(exp, "EXPERIMENT"))
```

## POLR on Survey Data Part I: Relatively more complex interactions

#### Ensure that survey data is ordered

```{r}
master$perts_q1msfixed1 <- ordered(master$perts_q1msfixed1, 
                                   levels=c("Strongly Disagree", "Disagree", 
                                            "Somewhat Disagree","Somewhat Agree", 
                                            "Agree", "Strongly Agree"))
master$perts_q1msfixed2 <- ordered(master$perts_q1msfixed2, 
                                   levels=c("Strongly Disagree", "Disagree", 
                                            "Somewhat Disagree", "Somewhat Agree", 
                                            "Agree", "Strongly Agree"))
```

### `perts_q1msfixed1` versus `perts_q1msfixed2` (`sex_id`)

```{r}
model_1 <- polr(perts_q1msfixed2 ~ exp*sex_id*perts_q1msfixed1 +
                  exp*sex_id + exp*perts_q1msfixed1 + sex_id*perts_q1msfixed1 +
                  exp + sex_id + perts_q1msfixed1 +
                  satm + satv + hs_gpa + aleksikc_score, data = master)

summary(model_1)

levels(master$perts_q1msfixed1)
```

The "L", "C", "Q", "4", and "5" suffixes on the `perts_q1msfixed` predictor variable are due to this predictor being an ordered factor. "L" stands for "linear", "C" stands for "cubic", and so on. R fits the data with a series of orthogonal polynomials (for some reason). I don't really know how to interpret the output, or why to ordering of the factor requires this sort of fitting.

If the predictor variables are *un-ordered* factors, then the summary output does not contain these suffixes. I observed this in my exploration of the `effects` package vingette (find it [here](https://uwnetid-my.sharepoint.com/personal/cfcraig_uw_edu/_layouts/15/onedrive.aspx?id=%2Fpersonal%2Fcfcraig%5Fuw%5Fedu%2Fresearch%2F0%20ANALYSES%2Fexplore%2Deffect%2Dpackage&listurl=%2Fpersonal%2Fcfcraig%5Fuw%5Fedu%2Fresearch%2F0%20ANALYSES&remoteItem=%7B%22mp%22%3A%7B%22webAbsoluteUrl%22%3A%22https%3A%2F%2Fuwnetid%2Dmy%2Esharepoint%2Ecom%2Fpersonal%2Fcfcraig%5Fuw%5Fedu%22%2C%22listFullUrl%22%3A%22https%3A%2F%2Fuwnetid%2Dmy%2Esharepoint%2Ecom%2Fpersonal%2Fcfcraig%5Fuw%5Fedu%2FDocuments%22%2C%22rootFolder%22%3A%22%2Fpersonal%2Fcfcraig%5Fuw%5Fedu%2FDocuments%2F0%20ANALYSES%22%7D%2C%22rsi%22%3A%7B%22listFullUrl%22%3A%22https%3A%2F%2Fuwnetid%2Dmy%2Esharepoint%2Ecom%2Fpersonal%2Fcfcraig%5Fuw%5Fedu%2Fresearch%2F0%20ANALYSES%22%2C%22rootFolder%22%3A%22%2Fpersonal%2Fcfcraig%5Fuw%5Fedu%2Fresearch%2F0%20ANALYSES%2Fexplore%2Deffect%2Dpackage%22%2C%22webAbsoluteUrl%22%3A%22https%3A%2F%2Fuwnetid%2Dmy%2Esharepoint%2Ecom%2Fpersonal%2Fcfcraig%5Fuw%5Fedu%2Fresearch%22%7D%7D&viewid=3a9f768c%2D0ade%2D43f5%2D8b58%2De85651c29766)).

This site helped me sort of understand what's going on with these suffixes: <https://data.library.virginia.edu/understanding-ordered-factors-in-a-linear-model/>

The `perts_q1msfixed` variable has six ordered levels, and the regression output shows five polynomial fits for `perts_q1msfixed`. However, these don't seem to have a 1:1 correspondence to the different levels of `perts_q1msfixed` . Is it more about about how many factor "units" different the outcome is relative to the input? That is, if a student's pre mindset response was "Disagree" and their post response was "Somewhat Disagree", is that a linear change because it differs by one factor unit? If the post response was "Somewhat Agree", is that a quadratic change?

**NOTES**:

-   Some of the mindset questions are coded backwards, so we'll have to address that!!
-   Don't I need to include "student" as a random effect? (Cynthia didn't do so in her survey analyses, from what I can tell.)
-   Did Cynthia ever look at individual survey scores?

```{r}
# This doesn't work. There is not a native way to include random effects in
# a polr model.
# model_1.1 <- polr(perts_q1msfixed2 ~ exp*sex_id*perts_q1msfixed1 +
#                   exp*sex_id + exp*perts_q1msfixed1 + sex_id*perts_q1msfixed1 +
#                   exp + sex_id + perts_q1msfixed1 +
#                   satm + satv + hs_gpa + aleksikc_score + (1|two_stage_id), 
#                   data = master)
# 
# summary(model_1.1)
```

Let's try the `effects` package:

```{r}
perts_q1_effects <- allEffects(model_1)
perts_q1_effects
```

polr(formula = perts_q1msfixed2 \~ exp \* sex_id \* perts_q1msfixed1 +

exp \* sex_id + exp \* perts_q1msfixed1 + sex_id \* perts_q1msfixed1 +

exp + sex_id + perts_q1msfixed1 + satm + satv + hs_gpa +

aleksikc_score, data = master)

```{r}
plot(effect("exp*sex_id*perts_q1msfixed1", model_1, 
     given.values = c(sex_idFemale = 0.5)))

plot(effect("exp*sex_id*perts_q1msfixed1", model_1, 
     given.values = c(sex_idFemale = 0.5)), style = "stacked")

plot(effect("exp*sex_id*perts_q1msfixed1", model_1, 
     given.values = c(sex_idFemale = 0.5), latent = TRUE))
```

```{r}
plot(effect("sex_id*perts_q1msfixed1", model_1, 
     given.values = c(sex_idFemale = 0.5)))

plot(effect("sex_id*perts_q1msfixed1", model_1, 
     given.values = c(sex_idFemale = 0.5)), style = "stacked")

plot(effect("sex_id*perts_q1msfixed1", model_1, 
     given.values = c(sex_idFemale = 0.5), latent = TRUE))
```

```{r}
plot(predictorEffect("satm", model_1, 
     given.values = c(sex_idFemale = 0.5)))
```

# Try some `anova` model selection:

```{r}
Anova(model_1)

model_1 <- polr(perts_q1msfixed2 ~ exp*sex_id*perts_q1msfixed1 +
                  exp*sex_id + exp*perts_q1msfixed1 + sex_id*perts_q1msfixed1 +
                  exp + sex_id + perts_q1msfixed1 +
                  satm + satv + hs_gpa + aleksikc_score, data = master)

summary(model_1)

```

**QUESITON**: How can a 3-way interaction be significant, but the related two-way interactions and main effects *not* be significant?

### `perts_q1msfixed1` versus `perts_q1msfixed2` (`urm_id`)

```{r}
model_2 <- polr(perts_q1msfixed2 ~ exp*urm_id*perts_q1msfixed1 +
                  exp*urm_id + exp*perts_q1msfixed1 + urm_id*perts_q1msfixed1 +
                  exp + urm_id + perts_q1msfixed1 +
                  satm + satv + hs_gpa + aleksikc_score, data = master)

summary(model_2)
```

### `perts_q1msfixed1` versus `perts_q1msfixed2` (`eop_id`)

```{r}
model_3 <- polr(perts_q1msfixed2 ~ exp*eop_id*perts_q1msfixed1 +
                  exp*eop_id + exp*perts_q1msfixed1 + eop_id*perts_q1msfixed1 +
                  exp + eop_id + perts_q1msfixed1 +
                  satm + satv + hs_gpa + aleksikc_score, data = master)

summary(model_3)
```

### `perts_q1msfixed1` versus `perts_q1msfixed2` (`fgn_id`)

```{r}
model_4 <- polr(perts_q1msfixed2 ~ exp*fgn_id*perts_q1msfixed1 +
                  exp*fgn_id + exp*perts_q1msfixed1 + fgn_id*perts_q1msfixed1 +
                  exp + fgn_id + perts_q1msfixed1 +
                  satm + satv + hs_gpa + aleksikc_score, data = master)

summary(model_4)
```

## POLR on Survey Data Part II: Relatively simpler interactions

#### Create a dataset that only contains data from the control year

```{r}
master_control <- master %>% 
  filter(exp == "CONTROL")
```

## `sex_id`:

### `perts_q1msfixed1` regressed on `exp` (`sex_id`; compared across years)

```{r}
model_5 <- polr(perts_q1msfixed2 ~ exp*sex_id +
                                    exp + sex_id +
                                    satm + satv + hs_gpa + aleksikc_score,
                                    data = master)

summary(model_5)
```

##### As the value of `exp` changes, the log odds of (males?) answering one or more higher levels on the Likert scale decreases by 0.111639

##### When we get rid of the other survey question as an interaction, the "L, Q, C, \^4 \^5" indicator in the regression output disappears

### `perts_q1msfixed1` versus `perts_q1msfixed2` (`sex_id`; compared within control year)

```{r}
model_6 <- polr(perts_q1msfixed2 ~ sex_id*perts_q1msfixed1 +
                                    sex_id + perts_q1msfixed1 +
                                    satm + satv + hs_gpa + aleksikc_score,
                                    data = master_control)

summary(model_6)
```

##### The log odds of females answering one or more higher levels on the Likert scale for the end of quarter survey vs. men decreases by 0.1677219?

## `urm_id`:

### `perts_q1msfixed1` regressed on `exp` (`urm_id`; compared across years)

```{r}
model_7 <- polr(perts_q1msfixed2 ~ exp*urm_id +
                                    exp + urm_id +
                                    satm + satv + hs_gpa + aleksikc_score,
                                    data = master)

summary(model_7)
```

### `perts_q1msfixed1` versus `perts_q1msfixed2` (`urm_id`; compared within control year)

```{r}
model_8 <- polr(perts_q1msfixed2 ~ urm_id*perts_q1msfixed1 +
                                    urm_id + perts_q1msfixed1 +
                                    satm + satv + hs_gpa + aleksikc_score,
                                    data = master_control)

summary(model_8)
```

## `eop_id`:

### `perts_q1msfixed1` regressed on `exp` (`eop_id`; compared across years)

```{r}
model_9 <- polr(perts_q1msfixed2 ~ exp*eop_id +
                                    exp + eop_id +
                                    satm + satv + hs_gpa + aleksikc_score,
                                    data = master)

summary(model_9)
```

### `perts_q1msfixed1` versus `perts_q1msfixed2` (`eop_id`; compared within control year)

```{r}
model_10 <- polr(perts_q1msfixed2 ~ eop_id*perts_q1msfixed1 +
                                    eop_id + perts_q1msfixed1 +
                                    satm + satv + hs_gpa + aleksikc_score,
                                    data = master_control)

summary(model_10)
```

## `fgn_id`:

### `perts_q1msfixed1` regressed on `exp` (`fgn_id`; compared across years)

```{r}
model_11 <- polr(perts_q1msfixed2 ~ exp*fgn_id +
                                    exp + fgn_id +
                                    satm + satv + hs_gpa + aleksikc_score,
                                    data = master)

summary(model_11)
```

### `perts_q1msfixed1` versus `perts_q1msfixed2` (`urm_id`; compared within control year)

```{r}
model_12 <- polr(perts_q1msfixed2 ~ fgn_id*perts_q1msfixed1 +
                                    fgn_id + perts_q1msfixed1 +
                                    satm + satv + hs_gpa + aleksikc_score,
                                    data = master_control)

summary(model_12)
```
