---
title: "Exploring POLR with Survey Data"
author: "Jackson Hughes"
date: "2/14/2022"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(skimr)
library(moderndive)
library(lme4)
library(MASS)
library(effects)
library(nnet)
library(Hmisc)
```

## Set up

### Designate working directories and copy file

```{r}
proj_dir <- here()
original_data_dir   <- here("original-data", "/")
importable_data_dir <- here("processing-and-analysis", "01-importable-data", "/")
analysis_data_dir   <- here("processing-and-analysis", "03-analysis-data", "/")
metadata_dir <- here("original-data", "metadata", "/")
copy_from <- paste0(original_data_dir, "two_stage_master_wide_deid.rds")
copy_to <- paste0(importable_data_dir, "two_stage_master_wide_deid.rds")
file.copy(copy_from, copy_to)
```

### Import dataset

```{r}
master_original_1 <- readRDS(paste0(importable_data_dir, "two_stage_master_wide_deid.rds"))
```

### Change value of NAs in `exp` column to `EXPERIMENT`

```{r}
master <- master_original_1 %>% 
  mutate(exp = replace_na(exp, "EXPERIMENT"))
```

## POLR on Survey Data

### `perts_q1msfixed1` versus `perts_q1msfixed2` (`sex_id`)

```{r}
master$perts_q1msfixed1 <- ordered(master$perts_q1msfixed1, levels=c("Strongly Disagree", "Disagree", "Somewhat Disagree", "Somewhat Agree", "Agree", "Strongly Agree"))
master$perts_q1msfixed2 <- ordered(master$perts_q1msfixed2, levels=c("Strongly Disagree", "Disagree", "Somewhat Disagree", "Somewhat Agree", "Agree", "Strongly Agree"))

model_1 <- polr(perts_q1msfixed2 ~ exp*sex_id*perts_q1msfixed1 +
                                    exp*sex_id + exp*perts_q1msfixed1 + sex_id*perts_q1msfixed1 +
                                    exp + sex_id + perts_q1msfixed1 +
                                    satm + satv + hs_gpa + aleksikc_score,
                                    data = master)

summary(model_1)
```

### `perts_q1msfixed1` versus `perts_q1msfixed2` (`urm_id`)

```{r}
model_2 <- polr(perts_q1msfixed2 ~ exp*urm_id*perts_q1msfixed1 +
                                    exp*urm_id + exp*perts_q1msfixed1 + urm_id*perts_q1msfixed1 +
                                    exp + urm_id + perts_q1msfixed1 +
                                    satm + satv + hs_gpa + aleksikc_score,
                                    data = master)

summary(model_2)