---
title: 'MLM: Random Intercepts Model'
author: "Jackson Hughes"
date: "5/24/2021"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(skimr)
library(moderndive)
library(lme4)
```

# Preliminary setup and data wrangling

## Data import

### Set working directories

```{r}
proj_dir <- here()
original_data_dir <- here("original-data", "/")
importable_data_dir <- here("processing-and-analysis", "01-importable-data", "/")
analysis_data_dir <- here("processing-and-analysis", "03-analysis-data", "/")
metadata_dir <- here("original-data", "metadata", "/")
```

### Copy data from `original-data` to `01-importable-data`

```{r}
copy_from <- paste0(original_data_dir, "master_2s_small_deid_scaled.rds")
copy_to <- paste0(importable_data_dir, "master_2s_small_deid_scaled.rds")
file.copy(copy_from, copy_to)
```

### Import master dataset

```{r}
master <- readRDS(paste0(importable_data_dir, "master_2s_small_deid_scaled.rds"))
```

## Data wrangling

### Condense dataset to only include unique rows; omit NAs

```{r}
master_extras_removed <- subset(master, select = -c(item_num, stud_ans, exam_key, corr, bloom_rating, complexity_rating_mean, complexity_rating_median, item_code, stem))
master_unique_1 <- unique(master_extras_removed, incomparables = FALSE)
master_unique_2 <- master_unique_1 %>% 
  select(exam1, exam2, finalexam, course_fullid, ta_sect, ver, sex_id, urm_id, eop_id, fgn_id, satverbal, satmath, high_sch_gpa, mastered_topics_initial_kc, experiment1, satm_c, satv_c, aleksikc_c, hsgpa_c, final_c, satm_z, satv_z, aleksikc_z, hsgpa_z, final_z)
master_unique <- na.omit(master_unique_2)
```

# Regression

## Random intercepts model for TA section (centered data)

```{r}
rand_int_mod_c <- lmer(final_c ~ experiment1 + satm_c + satv_c + aleksikc_c + hsgpa_c + (1|ta_sect), data = master_unique)
summary(rand_int_mod_c)
```

## Random intercepts model for TA section (z-scored data)

```{r}
rand_int_mod_z <- lmer(final_z ~ experiment1 + satm_z + satv_z + aleksikc_z + hsgpa_z + (1|ta_sect), data = master_unique)
summary(rand_int_mod_z)
```

## Notes
* Not sure how to interpret the summary data yet
* What is "Correlation of Fixed Effects"?